# =============================================================================
# 1_IMC-Hyperion_Seurat_clustering
#
# Description:
# This script performs single-cell protein clustering of Imaging Mass Cytometry
# (IMC) data using the Seurat framework. 
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - IMC-Hyperion_metadata.csv
# - Preprocessed IMC single-cell expression matrices (*.csv)
#
# Output:
# - Integrated Seurat object
# - PCA, Harmony, and UMAP embeddings
# - Cluster annotations
# - Visualization figures
#
# Notes:
# Analysis parameters are currently configured for the Precentral gyrus dataset.
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Create Seurat objects from CSV files
```{r}
metadata_file <- "IMC-Hyperion_metadata.csv"

SOc_metadata <- read.csv(
  metadata_file,
  stringsAsFactors = FALSE
)

# Select the precentral gyrus dataset
Data_selected <- SOc_metadata[
  SOc_metadata$Region == "PrecentralGyrus",
  ,
  drop = FALSE
]

# Create an input-file table
Data_table <- data.frame(
  path        = Data_selected$csv_R_analysis,
  group       = Data_selected$Name,
  sample_name = Data_selected$Sample_name,
  sex         = Data_selected$Sex,
  age         = Data_selected$Age,
  major_group = Data_selected$Group,
  stringsAsFactors = FALSE
)

# Convert each row into a list
files <- lapply(seq_len(nrow(Data_table)), function(i) {
  list(
    path        = Data_table$path[i],
    group       = Data_table$group[i],
    sample_name = Data_table$sample_name[i],
    sex         = Data_table$sex[i],
    age         = Data_table$age[i],
    major_group = Data_table$major_group[i]
  )
})

# Create a Seurat object for each ROI
seurat_list <- vector("list", length(files))

for (i in seq_along(files)) {

  # Load the single-cell protein-expression table
  temp <- read.csv(
    files[[i]]$path,
    header = TRUE,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  # Retain cells that passed cell-selection quality control
  seurat_data <- temp[
    temp$Sel01 %in% c(1, "1"),
    ,
    drop = FALSE
  ]

  # Select CellId and protein-marker columns
  seurat_data <- seurat_data[, 2:32, drop = FALSE]

  # Use CellId as the cell name
  rownames(seurat_data) <- seurat_data$CellId
  seurat_data <- seurat_data[, 2:31, drop = FALSE]

  # Convert cells × markers into markers × cells
  expression_matrix <- t(as.matrix(seurat_data))
  storage.mode(expression_matrix) <- "numeric"

  # Create a Seurat object
  seurat_object <- CreateSeuratObject(
    counts = expression_matrix,
    assay = "Protein"
  )

  DefaultAssay(seurat_object) <- "Protein"

  # Add sample metadata
  seurat_object$group       <- files[[i]]$group
  seurat_object$sample_name <- files[[i]]$sample_name
  seurat_object$sex         <- files[[i]]$sex
  seurat_object$age         <- files[[i]]$age
  seurat_object$major_group <- files[[i]]$major_group

  # CLR normalization and scaling
  seurat_object <- NormalizeData(
    seurat_object,
    assay = "Protein",
    normalization.method = "CLR",
    margin = 2
  )

  seurat_object <- ScaleData(
    seurat_object,
    assay = "Protein"
  )

  seurat_list[[i]] <- seurat_object
}

# Group ROI-level Seurat objects by sample
sample_names <- vapply(
  seurat_list,
  function(x) as.character(unique(x$sample_name)),
  character(1)
)

by_sample <- split(
  seurat_list,
  sample_names
)

# Assign an ROI identifier within each sample
for (sample_id in names(by_sample)) {

  roi_list <- by_sample[[sample_id]]

  for (j in seq_along(roi_list)) {
    roi_list[[j]]$roi <- paste0(sample_id, "_ROI", j)
  }

  by_sample[[sample_id]] <- roi_list
}

# Merge ROIs within each sample
seurat_per_sample <- lapply(by_sample, function(roi_list) {

  merged_object <- Reduce(
    function(x, y) merge(x, y),
    roi_list
  )

  # Join assay layers when using Seurat v5
  if ("layers" %in% slotNames(merged_object[["Protein"]])) {
    merged_object[["Protein"]] <-
      JoinLayers(merged_object[["Protein"]])
  }

  merged_object
})

# Merge all samples
combined_seurat <- Reduce(
  function(x, y) merge(x, y),
  seurat_per_sample
)

# Join the merged assay layers
if ("layers" %in% slotNames(combined_seurat[["Protein"]])) {
  combined_seurat[["Protein"]] <-
    JoinLayers(combined_seurat[["Protein"]])
}

# Initially include all measured protein markers
VariableFeatures(combined_seurat) <- rownames(combined_seurat)

# Markers excluded from dimensionality reduction
excluded_markers <- c(
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

pca_features <- setdiff(
  VariableFeatures(combined_seurat),
  excluded_markers
)

VariableFeatures(combined_seurat) <- pca_features

# PCA
combined_seurat <- RunPCA(
  combined_seurat,
  assay = "Protein",
  features = VariableFeatures(combined_seurat)
)
```

## Batch Effect Correction
```{r}
combined_seurat <- RunHarmony(
  object = combined_seurat,
   group.by.vars = c("roi")
)
```

## UMAP clustring
Note: UMAP embeddings may show slight differences in layout across runs owing to stochastic optimization. For exact reproduction of previously generated embeddings, the saved RDS object should be loaded instead of rerunning the workflow.
```{r}
combined_seurat <- FindNeighbors(combined_seurat,
                                 reduction = "harmony",
                                 dims      = 1:17,
                                 k.param   = 50)  

combined_seurat <- FindClusters(combined_seurat, resolution = 1.4)

combined_seurat <- RunUMAP(combined_seurat,
                           reduction = "harmony",
                           dims      = 1:17,
                           min.dist  = 0.3,
                           spread    = 3,
                           n.neighbors = 30) 

DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, group.by = "run_id")
DimPlot(combined_seurat, label = T, group.by = "group")
DimPlot(combined_seurat, label = T, group.by = "Sample_name")
DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, split.by = "group")
DimPlot(combined_seurat, label = T, split.by = "Sample_name")
```

## Heatmap to check marker expression patterns
```{r}
DefaultLayer(combined_seurat[["Protein"]]) <- "data"

markers_astrocyte <- c(
  "Cell-141Pr-AQP4",
  "Cell-158Gd-GFAP"
)

markers_immune <- c(
  "Cell-161Dy-CX3CR1",
  "Cell-164Dy-TMEM119",
  "Cell-160Gd-IBA1",
  "Cell-163Dy-CD4",
  "Cell-155Gd-FoxP3",
  "Cell-167Er-CD8"
)

markers_neuron <- c(
  "Cell-151Eu-MAP2",
  "Cell-149Sm-C9orf72", # NeuN
  "Cell-169Tm-ChAT",
  "Cell-166Er-GAD67"
)

markers_others <- c(
  "Cell-175Lu-TDP43",
  "Cell-154Sm-pTDP43",
  "Cell-168Er-FUS",
  "Cell-152Sm-ATXN2",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-176Yb-TBK1",
  "Cell-162Dy-pTBK1",
  "Cell-147Sm-Glutaminase",
  "Cell-143Nd-SQSTM1",
  "Cell-156Gd-CH24H",
  "Cell-153Eu-ORF1",
  "Cell-159Tb-IL6ST",
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

markers <- c(
  markers_astrocyte,
  markers_immune,
  markers_neuron,
  markers_others
)

marker_labels <- c(
  "Cell-141Pr-AQP4"          = "AQP4",
  "Cell-143Nd-SQSTM1"        = "SQSTM1",
  "Cell-145Nd-UBQLN2"        = "UBQLN2",
  "Cell-147Sm-Glutaminase"   = "GLS",
  "Cell-149Sm-C9orf72"       = "NeuN",
  "Cell-150Nd-PDL1"          = "PDL1",
  "Cell-151Eu-MAP2"          = "MAP2",
  "Cell-152Sm-ATXN2"         = "ATXN2",
  "Cell-153Eu-ORF1"          = "ORF1",
  "Cell-154Sm-pTDP43"        = "pTDP43",
  "Cell-155Gd-FoxP3"         = "FOXP3",
  "Cell-156Gd-CH24H"         = "CH24H",
  "Cell-158Gd-GFAP"          = "GFAP",
  "Cell-159Tb-IL6ST"         = "IL6ST",
  "Cell-160Gd-IBA1"          = "IBA1",
  "Cell-161Dy-CX3CR1"        = "CX3CR1",
  "Cell-162Dy-pTBK1"         = "pTBK1",
  "Cell-163Dy-CD4"           = "CD4",
  "Cell-164Dy-TMEM119"       = "TMEM119",
  "Cell-165Ho-PD1"           = "PD1",
  "Cell-166Er-GAD67"         = "GAD67",
  "Cell-167Er-CD8"           = "CD8",
  "Cell-168Er-FUS"           = "FUS",
  "Cell-169Tm-ChAT"          = "ChAT",
  "Cell-170Er-APOE"          = "APOE",
  "Cell-171Yb-TREM2"         = "TREM2",
  "Cell-172Yb-PDL2"          = "PDL2",
  "Cell-174Yb-MHCclassII"    = "MHCII",
  "Cell-175Lu-TDP43"         = "TDP43",
  "Cell-176Yb-TBK1"          = "TBK1"
)

heatmap_data <- DotPlot(combined_seurat, features = markers)$data
heatmap_data$id <- factor(
  heatmap_data$id,
  levels = sort(as.numeric(unique(heatmap_data$id)))
)

p <- ggplot(heatmap_data, aes(x = features.plot, y = fct_rev(id))) +
  geom_tile(colour = "white", aes(fill = avg.exp.scaled)) +
  scale_fill_gradientn(
    colours = c("lightyellow3", "lightyellow3", "lightyellow", "orange", "red"),
    values = c(0, 0.2, 0.5, 0.75, 1),
    name = "Normalized Expression"
  ) +
  scale_x_discrete(labels = marker_label) +
  ggtitle("Heatmap of marker expression") +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, face = "italic"),
    legend.position = "bottom"
  )

print(p)
```

## Major cell-type annotation
```{r}
cluster_to_celltype <- c(
  "0"  = "Excitatory",
  "1"  = "Excitatory",
  "2"  = "Others",
  "3"  = "Others",
  "4"  = "Inhibitory",
  "5"  = "Astrocyte",
  "6"  = "Excitatory",
  "7"  = "Others",
  "8"  = "Astrocyte",
  "9"  = "Excitatory",
  "10" = "Immune",
  "11" = "Excitatory",
  "12" = "Immune"
)

annotated_object <- RenameIdents(
  combined_seurat,
  cluster_to_celltype
)

combined_seurat$Celltype <- Idents(annotated_object)

combined_seurat$Celltype <- factor(
  combined_seurat$Celltype,
  levels = c(
    "Astrocyte",
    "Immune",
    "Excitatory",
    "Inhibitory",
    "Others"
  )
)

celltype_colors <- c(
  "Astrocyte"  = "#F8766D",
  "Immune"     = "#FFD700",
  "Excitatory" = "#00B0F6",
  "Inhibitory" = "#00BF7D",
  "Others"     = "#A3A500"
)

p <- DimPlot(
  combined_seurat,
  group.by = "Celltype",
  pt.size = 0.3,
  label = TRUE,
  label.size = 6
) +
  scale_color_manual(
    values = celltype_colors
  ) +
  labs(
    title = "Major cell type",
    color = "Cell type"
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(
      size = 20,
      color = "black"
    ),
    axis.title = element_text(
      size = 20,
      color = "black"
    ),
    legend.position = "right",
    axis.line = element_line(
      color = "black",
      linewidth = 0.5
    ),
    axis.ticks.length = unit(5, "pt")
  )

print(p)

ggsave(
  filename = file.path(output_folder, "UMAP_CellType.png"),
  plot = p,
  width = 8,
  height = 7,
  dpi = 600,
  bg = "white"
)
```

## Save integrated Seurat object
```{r}
#saveRDS(
#  combined_seurat,
#  file = file.path(output_folder, "combined_seurat_PrecentralGyrus.rds")
#)
```


