# =============================================================================
# 2_IMC-Hyperion_Heatmap.Rmd
#
# Description:
# This script calculates mean protein expression for each major cell type,
# Z-score-scales the mean expression across cell types for each marker,
# and generates a heatmap.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object 
#
# Output:
# - Heatmap of Z-scored mean protein expression
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load the integrated Seurat object
```{r}
combined_seurat<-readRDS("combined_seurat_PrecentralGyrus.rds")
```

## Define marker panels
```{r}
markers_astrocyte <- c(
  "Cell-141Pr-AQP4",
  "Cell-158Gd-GFAP"
)

markers_immune <- c(
  "Cell-161Dy-CX3CR1",
  "Cell-164Dy-TMEM119",
  "Cell-160Gd-IBA1",
  "Cell-163Dy-CD4",
  "Cell-155Gd-FoxP3",
  "Cell-167Er-CD8"
)

markers_neuron <- c(
  "Cell-151Eu-MAP2",
  "Cell-149Sm-C9orf72", # NeuN
  "Cell-169Tm-ChAT",
  "Cell-166Er-GAD67"
)

markers_others <- c(
  "Cell-175Lu-TDP43",
  "Cell-154Sm-pTDP43",
  "Cell-168Er-FUS",
  "Cell-152Sm-ATXN2",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-176Yb-TBK1",
  "Cell-162Dy-pTBK1",
  "Cell-147Sm-Glutaminase",
  "Cell-143Nd-SQSTM1",
  "Cell-156Gd-CH24H",
  "Cell-153Eu-ORF1",
  "Cell-159Tb-IL6ST",
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

markers <- c(
  markers_astrocyte,
  markers_immune,
  markers_neuron,
  markers_others
)

marker_labels <- c(
  "Cell-141Pr-AQP4"          = "AQP4",
  "Cell-143Nd-SQSTM1"        = "SQSTM1",
  "Cell-145Nd-UBQLN2"        = "UBQLN2",
  "Cell-147Sm-Glutaminase"   = "GLS",
  "Cell-149Sm-C9orf72"       = "NeuN",
  "Cell-150Nd-PDL1"          = "PDL1",
  "Cell-151Eu-MAP2"          = "MAP2",
  "Cell-152Sm-ATXN2"         = "ATXN2",
  "Cell-153Eu-ORF1"          = "ORF1",
  "Cell-154Sm-pTDP43"        = "pTDP43",
  "Cell-155Gd-FoxP3"         = "FOXP3",
  "Cell-156Gd-CH24H"         = "CH24H",
  "Cell-158Gd-GFAP"          = "GFAP",
  "Cell-159Tb-IL6ST"         = "IL6ST",
  "Cell-160Gd-IBA1"          = "IBA1",
  "Cell-161Dy-CX3CR1"        = "CX3CR1",
  "Cell-162Dy-pTBK1"         = "pTBK1",
  "Cell-163Dy-CD4"           = "CD4",
  "Cell-164Dy-TMEM119"       = "TMEM119",
  "Cell-165Ho-PD1"           = "PD1",
  "Cell-166Er-GAD67"         = "GAD67",
  "Cell-167Er-CD8"           = "CD8",
  "Cell-168Er-FUS"           = "FUS",
  "Cell-169Tm-ChAT"          = "ChAT",
  "Cell-170Er-APOE"          = "APOE",
  "Cell-171Yb-TREM2"         = "TREM2",
  "Cell-172Yb-PDL2"          = "PDL2",
  "Cell-174Yb-MHCclassII"    = "MHCII",
  "Cell-175Lu-TDP43"         = "TDP43",
  "Cell-176Yb-TBK1"          = "TBK1"
)
```

## Define the cell-type order
```{r}
celltype_order <- c(
  "Astrocyte",
  "Immune",
  "Excitatory",
  "Inhibitory",
  "Others"
)

combined_seurat$Celltype <- factor(
  combined_seurat$Celltype,
  levels = celltype_order
)
```


## Calculate mean and Z-scored expression
```{r}
make_heatmap_data <- function(
    seu,
    group_var,
    markers,
    marker_labels,
    clip_range = c(-2.5, 2.5)
) {

  # Reverse the marker order for bottom-to-top heatmap display
  markers_rev <- rev(markers)

  # Extract normalized protein-expression values
  expression_data <- FetchData(
    seu,
    vars = markers_rev,
    layer = "data"
  )

  expression_data$Group <- seu[[group_var, drop = TRUE]]

  # Calculate mean expression within each cell type
  mean_data <- expression_data %>%
    group_by(Group) %>%
    summarise(
      across(
        all_of(markers_rev),
        ~ mean(.x, na.rm = TRUE)
      ),
      .groups = "drop"
    )

  # Convert to long format and Z-score each marker across cell types
  heatmap_data <- mean_data %>%
    pivot_longer(
      cols = all_of(markers_rev),
      names_to = "marker",
      values_to = "mean_expression"
    ) %>%
    group_by(marker) %>%
    mutate(
      z_score = as.numeric(scale(mean_expression)),
      z_score = scales::squish(
        z_score,
        range = clip_range
      )
    ) %>%
    ungroup()

  # Set marker labels and display order
  heatmap_data$marker_label <- factor(
    heatmap_data$marker,
    levels = markers_rev,
    labels = marker_labels[markers_rev]
  )

  heatmap_data
}
```


## Heatmap function
```{r}
plot_heatmap <- function(
    heatmap_data,
    group_levels,
    clip_range = c(-2.5, 2.5)
) {

  heatmap_data$Group <- factor(
    heatmap_data$Group,
    levels = group_levels
  )

  ggplot(
    heatmap_data,
    aes(
      x = Group,
      y = marker_label,
      fill = z_score
    )
  ) +
    geom_tile(
      colour = "white",
      linewidth = 0.4
    ) +
    scale_fill_gradientn(
      colours = c(
        "lightyellow3",
        "lightyellow",
        "orange",
        "red"
      ),
      values = scales::rescale(
        c(clip_range[1], 0, 1, clip_range[2])
      ),
      limits = clip_range,
      oob = scales::squish,
      name = "Z-scored\nexpression",
      guide = guide_colorbar(
        title.theme = element_text(
          color = "black",
          size = 16
        ),
        label.theme = element_text(
          color = "black",
          size = 12
        )
      )
    ) +
    scale_x_discrete(
      position = "top",
      drop = FALSE
    ) +
    theme_classic(
      base_family = "Arial"
    ) +
    theme(
      axis.text.x = element_text(
        angle = 80,
        size = 18,
        color = "black",
        vjust = 1,
        hjust = 0
      ),
      axis.text.y = element_text(
        size = 18,
        color = "black"
      ),
      axis.title = element_blank(),
      legend.position = "bottom",
      axis.line = element_line(
        color = "black",
        linewidth = 0.5
      ),
      axis.ticks = element_line(
        color = "black",
        linewidth = 0.5
      ),
      axis.ticks.length = grid::unit(
        5,
        "pt"
      )
    )
}
```


## Generate the heatmap data and plot
```{r}
heatmap_data <- make_heatmap_data(
  seu = combined_seurat,
  group_var = "Celltype",
  markers = markers,
  marker_labels = marker_labels
)

p <- plot_heatmap(
  heatmap_data = heatmap_data,
  group_levels = celltype_order
)

print(p)

ggsave(
  filename = file.path(
    output_folder,
    "Heatmap_Celltype.png"
  ),
  plot = p,
  width = 3.5,
  height = 15,
  units = "in",
  dpi = 300,
  limitsize = FALSE,
  bg = "white"
)
```

