# =============================================================================
# 3_IMC-Hyperion_DEPs_dotplot.Rmd
#
# Description:
# This script computes differential marker expression between ALS and control
# groups within each major cell type from the integrated Imaging Mass
# Cytometry (IMC) Seurat object and visualizes the results as a dotplot.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object 
#
# Output:
# - Dotplot showing differential marker expression across cell types
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

# Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(ggplot2)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load the integrated Seurat object
```{r}
combined_seurat<-readRDS("combined_seurat_PrecentralGyrus.rds")
```

## Define marker panels
```{r}
markers_astrocyte <- c(
  "Cell-141Pr-AQP4",
  "Cell-158Gd-GFAP"
)

markers_immune <- c(
  "Cell-161Dy-CX3CR1",
  "Cell-164Dy-TMEM119",
  "Cell-160Gd-IBA1",
  "Cell-163Dy-CD4",
  "Cell-155Gd-FoxP3",
  "Cell-167Er-CD8"
)

markers_neuron <- c(
  "Cell-151Eu-MAP2",
  "Cell-149Sm-C9orf72", # NeuN
  "Cell-169Tm-ChAT",
  "Cell-166Er-GAD67"
)

markers_others <- c(
  "Cell-175Lu-TDP43",
  "Cell-154Sm-pTDP43",
  "Cell-168Er-FUS",
  "Cell-152Sm-ATXN2",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-176Yb-TBK1",
  "Cell-162Dy-pTBK1",
  "Cell-147Sm-Glutaminase",
  "Cell-143Nd-SQSTM1",
  "Cell-156Gd-CH24H",
  "Cell-153Eu-ORF1",
  "Cell-159Tb-IL6ST",
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

markers <- c(
  markers_astrocyte,
  markers_immune,
  markers_neuron,
  markers_others
)

marker_labels <- c(
  "Cell-141Pr-AQP4"          = "AQP4",
  "Cell-143Nd-SQSTM1"        = "SQSTM1",
  "Cell-145Nd-UBQLN2"        = "UBQLN2",
  "Cell-147Sm-Glutaminase"   = "GLS",
  "Cell-149Sm-C9orf72"       = "NeuN",
  "Cell-150Nd-PDL1"          = "PDL1",
  "Cell-151Eu-MAP2"          = "MAP2",
  "Cell-152Sm-ATXN2"         = "ATXN2",
  "Cell-153Eu-ORF1"          = "ORF1",
  "Cell-154Sm-pTDP43"        = "pTDP43",
  "Cell-155Gd-FoxP3"         = "FOXP3",
  "Cell-156Gd-CH24H"         = "CH24H",
  "Cell-158Gd-GFAP"          = "GFAP",
  "Cell-159Tb-IL6ST"         = "IL6ST",
  "Cell-160Gd-IBA1"          = "IBA1",
  "Cell-161Dy-CX3CR1"        = "CX3CR1",
  "Cell-162Dy-pTBK1"         = "pTBK1",
  "Cell-163Dy-CD4"           = "CD4",
  "Cell-164Dy-TMEM119"       = "TMEM119",
  "Cell-165Ho-PD1"           = "PD1",
  "Cell-166Er-GAD67"         = "GAD67",
  "Cell-167Er-CD8"           = "CD8",
  "Cell-168Er-FUS"           = "FUS",
  "Cell-169Tm-ChAT"          = "ChAT",
  "Cell-170Er-APOE"          = "APOE",
  "Cell-171Yb-TREM2"         = "TREM2",
  "Cell-172Yb-PDL2"          = "PDL2",
  "Cell-174Yb-MHCclassII"    = "MHCII",
  "Cell-175Lu-TDP43"         = "TDP43",
  "Cell-176Yb-TBK1"          = "TBK1"
)
```

## Define the analysis settings
```{r}
celltype_order <- c(
  "Astrocyte",
  "Immune",
  "Excitatory",
  "Inhibitory",
  "Others"
)

# Retain only cell types present in the Seurat object
celltypes_use <- celltype_order[
  celltype_order %in% unique(combined_seurat$Celltype)
]

# Markers displayed in the differential-expression dot plot
selected_markers <- markers_others

# Display ranges
cohens_d_range <- c(-1.5, 1.5)
q_threshold <- 10
```


## Calculate differential marker expression
```{r}
compute_stats_one <- function(
    seu,
    celltype,
    markers,
    assay = "Protein",
    layer = "counts",
    q_threshold = 10
) {

  # Select ALS and control cells belonging to the specified cell type
  object <- subset(
    seu,
    subset =
      Celltype == celltype &
      major_group %in% c("ALS", "Control")
  )

  # Extract protein-expression values
  expression_matrix <- as.matrix(
    GetAssayData(
      object,
      assay = assay,
      layer = layer
    )
  )

  # Inverse hyperbolic sine transformation
  expression_matrix <- asinh(expression_matrix)

  als_cells <- colnames(object)[
    object$major_group == "ALS"
  ]

  control_cells <- colnames(object)[
    object$major_group == "Control"
  ]

  if (length(als_cells) < 2 || length(control_cells) < 2) {
    warning(
      "Insufficient cells for ",
      celltype,
      "; this cell type was skipped."
    )
    return(NULL)
  }

  # Group means
  mean_als <- rowMeans(
    expression_matrix[, als_cells, drop = FALSE],
    na.rm = TRUE
  )

  mean_control <- rowMeans(
    expression_matrix[, control_cells, drop = FALSE],
    na.rm = TRUE
  )

  mean_all <- rowMeans(
    expression_matrix,
    na.rm = TRUE
  )

  # Group standard deviations
  sd_als <- apply(
    expression_matrix[, als_cells, drop = FALSE],
    1,
    sd,
    na.rm = TRUE
  )

  sd_control <- apply(
    expression_matrix[, control_cells, drop = FALSE],
    1,
    sd,
    na.rm = TRUE
  )

  n_als <- length(als_cells)
  n_control <- length(control_cells)

  # Pooled standard deviation
  pooled_sd <- sqrt(
    (
      (n_als - 1) * sd_als^2 +
      (n_control - 1) * sd_control^2
    ) /
      (n_als + n_control - 2)
  )

  # Cohen's d
  cohens_d <- ifelse(
    pooled_sd > 0,
    (mean_als - mean_control) / pooled_sd,
    NA_real_
  )

  # Log2 fold change calculated from the transformed mean values
  log2fc <- log2(
    (mean_als + 1) /
      (mean_control + 1)
  )

  # Wilcoxon rank-sum test
  p_values <- apply(
    expression_matrix,
    1,
    function(expression_values) {
      wilcox.test(
        expression_values[als_cells],
        expression_values[control_cells],
        exact = FALSE
      )$p.value
    }
  )

  # Benjamini-Hochberg correction across all measured markers
  # within the current cell type
  p_adjusted <- p.adjust(
    p_values,
    method = "BH"
  )

  minus_log10_fdr <- -log10(p_adjusted)

  # Custom composite score
  q_value <- abs(cohens_d) * minus_log10_fdr

  results <- data.frame(
    Marker = rownames(expression_matrix),
    Celltype = celltype,
    n_ALS = n_als,
    n_Control = n_control,
    Mean_ALS = mean_als,
    Mean_Control = mean_control,
    Mean_all = mean_all,
    log2FC = log2fc,
    Cohens_d = cohens_d,
    P_value = p_values,
    FDR = p_adjusted,
    minus_log10_FDR = minus_log10_fdr,
    q_value = q_value,
    q_above_threshold = q_value > q_threshold,
    stringsAsFactors = FALSE
  )

  # Retain the markers selected for visualization
  results <- results[
    results$Marker %in% markers,
    ,
    drop = FALSE
  ]

  results
}
```

## Apply the analysis to each major cell type
```{r}
stats_list <- lapply(
  celltypes_use,
  compute_stats_one,
  seu = combined_seurat,
  markers = selected_markers,
  assay = "Protein",
  layer = "counts",
  q_threshold = q_threshold
)

stats_all <- bind_rows(stats_list)
```


## Set the marker and cell-type order
```{r}
marker_order <- selected_markers[
  selected_markers %in% stats_all$Marker
]

stats_all$Marker <- factor(
  stats_all$Marker,
  levels = rev(marker_order)
)

stats_all$Celltype <- factor(
  stats_all$Celltype,
  levels = celltypes_use
)
```


## Scale the mean expression used for dot size
```{r}
stats_all <- stats_all %>%
  group_by(Marker) %>%
  mutate(
    Mean_expression_scaled = {
      quantile_limits <- quantile(
        Mean_all,
        probs = c(0.01, 0.95),
        na.rm = TRUE
      )

      clipped_values <- scales::squish(
        Mean_all,
        range = quantile_limits
      )

      if (diff(range(clipped_values, na.rm = TRUE)) == 0) {
        rep(0.5, length(clipped_values))
      } else {
        scales::rescale(
          clipped_values,
          to = c(0, 1)
        )
      }
    }
  ) %>%
  ungroup()

stats_all$Significance <- ifelse(
  stats_all$q_above_threshold,
  "*",
  ""
)
```


## Generate the differential-expression dot plot
```{r}
p <- ggplot(
  stats_all,
  aes(
    x = Celltype,
    y = Marker
  )
) +
  geom_point(
    aes(
      size = Mean_expression_scaled,
      fill = Cohens_d
    ),
    shape = 21,
    colour = "grey40",
    stroke = 0.5
  ) +
  geom_text(
    aes(label = Significance),
    size = 12,
    vjust = 0.8
  ) +
  scale_size(
    range = c(2, 15),
    limits = c(0, 1),
    name = "Relative\nexpression"
  ) +
  scale_fill_gradientn(
    colours = c(
      "blue",
      "#F2F2F2",
      "white",
      "#F2F2F2",
      "red"
    ),
    values = scales::rescale(
      c(-1.5, -0.2, 0, 0.2, 1.5)
    ),
    limits = cohens_d_range,
    oob = scales::squish,
    name = "Cohen's d"
  ) +
  scale_y_discrete(
    labels = marker_labels
  ) +
  theme_classic(
    base_family = "Arial"
  ) +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_text(
      size = 16,
      color = "black"
    ),
    axis.text.x = element_text(
      angle = 90,
      size = 16,
      color = "black",
      vjust = 0.5,
      hjust = 1
    ),
    legend.title = element_text(
      size = 12
    ),
    legend.text = element_text(
      size = 10
    )
  )

print(p)

ggsave(
  filename = file.path(
    output_folder,
    "DEP_dotplot.png"
  ),
  plot = p,
  width = 6,
  height = 10,
  units = "in",
  dpi = 300,
  limitsize = FALSE,
  bg = "white"
)
```
