# =============================================================================
# 4_IMC-Hyperion_Reclustering.Rmd
#
# Description:
# This script reclusters astrocytes, excitatory neurons, and immune cells from
# the precentral gyrus IMC dataset and visualizes the resulting subcell types.
# It also generates a marker-expression dot plot and heatmap.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object 
#
# Output:
# - Reclustering UMAPs for astrocytes, excitatory neurons, and immune cells
# - Combined marker-expression dot plot
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

# Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(patchwork)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load the integrated Seurat object
```{r}
combined_seurat<-readRDS("combined_seurat_PrecentralGyrus.rds")
```

## Define marker panels
```{r}
markers_astrocyte <- c(
  "Cell-141Pr-AQP4",
  "Cell-158Gd-GFAP"
)

markers_immune <- c(
  "Cell-161Dy-CX3CR1",
  "Cell-164Dy-TMEM119",
  "Cell-160Gd-IBA1",
  "Cell-163Dy-CD4",
  "Cell-155Gd-FoxP3",
  "Cell-167Er-CD8"
)

markers_neuron <- c(
  "Cell-151Eu-MAP2",
  "Cell-149Sm-C9orf72", # NeuN
  "Cell-169Tm-ChAT",
  "Cell-166Er-GAD67"
)

markers_others <- c(
  "Cell-175Lu-TDP43",
  "Cell-154Sm-pTDP43",
  "Cell-168Er-FUS",
  "Cell-152Sm-ATXN2",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-176Yb-TBK1",
  "Cell-162Dy-pTBK1",
  "Cell-147Sm-Glutaminase",
  "Cell-143Nd-SQSTM1",
  "Cell-156Gd-CH24H",
  "Cell-153Eu-ORF1",
  "Cell-159Tb-IL6ST",
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

markers <- c(
  markers_astrocyte,
  markers_immune,
  markers_neuron,
  markers_others
)

marker_labels <- c(
  "Cell-141Pr-AQP4"          = "AQP4",
  "Cell-143Nd-SQSTM1"        = "SQSTM1",
  "Cell-145Nd-UBQLN2"        = "UBQLN2",
  "Cell-147Sm-Glutaminase"   = "GLS",
  "Cell-149Sm-C9orf72"       = "NeuN",
  "Cell-150Nd-PDL1"          = "PDL1",
  "Cell-151Eu-MAP2"          = "MAP2",
  "Cell-152Sm-ATXN2"         = "ATXN2",
  "Cell-153Eu-ORF1"          = "ORF1",
  "Cell-154Sm-pTDP43"        = "pTDP43",
  "Cell-155Gd-FoxP3"         = "FOXP3",
  "Cell-156Gd-CH24H"         = "CH24H",
  "Cell-158Gd-GFAP"          = "GFAP",
  "Cell-159Tb-IL6ST"         = "IL6ST",
  "Cell-160Gd-IBA1"          = "IBA1",
  "Cell-161Dy-CX3CR1"        = "CX3CR1",
  "Cell-162Dy-pTBK1"         = "pTBK1",
  "Cell-163Dy-CD4"           = "CD4",
  "Cell-164Dy-TMEM119"       = "TMEM119",
  "Cell-165Ho-PD1"           = "PD1",
  "Cell-166Er-GAD67"         = "GAD67",
  "Cell-167Er-CD8"           = "CD8",
  "Cell-168Er-FUS"           = "FUS",
  "Cell-169Tm-ChAT"          = "ChAT",
  "Cell-170Er-APOE"          = "APOE",
  "Cell-171Yb-TREM2"         = "TREM2",
  "Cell-172Yb-PDL2"          = "PDL2",
  "Cell-174Yb-MHCclassII"    = "MHCII",
  "Cell-175Lu-TDP43"         = "TDP43",
  "Cell-176Yb-TBK1"          = "TBK1"
)
```

## Reclustering function
```{r}
recluster_celltype <- function(
    seu,
    celltype,
    dims,
    k_param,
    resolution,
    n_neighbors,
    min_dist,
    spread,
    seed,
    umap_seed = seed
) {
  object <- subset(
    seu,
    subset = Celltype == celltype
  )

  DefaultAssay(object) <- "Protein"
  VariableFeatures(object) <- rownames(object)

  object <- RunPCA(
    object,
    assay = "Protein",
    features = VariableFeatures(object),
    verbose = FALSE
  )

  set.seed(seed)

  object <- FindNeighbors(
    object,
    reduction = "pca",
    dims = dims,
    k.param = k_param,
    verbose = FALSE
  )

  object <- FindClusters(
    object,
    resolution = resolution,
    verbose = FALSE
  )

  object <- RunUMAP(
    object,
    reduction = "pca",
    dims = dims,
    n.neighbors = n_neighbors,
    min.dist = min_dist,
    spread = spread,
    seed.use = umap_seed,
    verbose = FALSE
  )

  object
}
```

## Recluster astrocytes
```{r}
combined_seurat_Astro <- recluster_celltype(
  seu = combined_seurat,
  celltype = "Astrocyte",
  prefix = "Astro_",
  dims = 1:6,
  k_param = 30,
  resolution = 0.9,
  n_neighbors = 30,
  min_dist = 0.2,
  spread = 2,
  seed = 11,
  umap_seed = 11
)

combined_seurat_Astro$Subcelltype <- factor(
  paste0(
    "Astro_",
    as.integer(as.character(Idents(combined_seurat_Astro))) + 1
  ),
  levels = paste0("Astro_", 1:5)
)
```

## Recluster excitatory neurons
```{r}
combined_seurat_ExcNeuron <- recluster_celltype(
  seu = combined_seurat,
  celltype = "Excitatory",
  prefix = "Excit_",
  dims = 1:6,
  k_param = 50,
  resolution = 0.6,
  n_neighbors = 50,
  min_dist = 0.1,
  spread = 2,
  seed = 11,
  umap_seed = 11
)

combined_seurat_ExcNeuron$Subcelltype <- factor(
  paste0(
    "Excit_",
    as.integer(as.character(Idents(combined_seurat_ExcNeuron))) + 1
  ),
  levels = paste0("Excit_", 1:4)
)
```

## Recluster immune cells
```{r}
combined_seurat_Immune <- recluster_celltype(
  seu = combined_seurat,
  celltype = "Immune",
  prefix = "Immune_",
  dims = 1:6,
  k_param = 20,
  resolution = 0.3,
  n_neighbors = 50,
  min_dist = 0.1,
  spread = 2,
  seed = 20,
  umap_seed = 15
)

combined_seurat_Immune$Subcelltype <- factor(
  paste0(
    "Immune_",
    as.integer(as.character(Idents(combined_seurat_Immune))) + 1
  ),
  levels = paste0("Immune_", 1:3)
)
```

Note: UMAP embeddings may show slight differences in layout across runs owing to stochastic optimization. For exact reproduction of previously generated embeddings, the saved RDS object should be loaded instead of rerunning the workflow.
## Reload the saved object for exact figure reproduction
```{r}

combined_seurat_Astro <- readRDS(
  "combined_seurat_PrecentralGyrus_Astrocyte.rds"
)

combined_seurat_ExcNeuron <- readRDS(
  "combined_seurat_PrecentralGyrus_ExcNeuron.rds"
)

combined_seurat_Immune <- readRDS(
  "combined_seurat_PrecentralGyrus_Immune.rds"
)

combined_seurat_Astro$Subcelltype <- factor(
  combined_seurat_Astro$Subcelltype,
  levels = paste0("Astro_", 1:5)
)

combined_seurat_ExcNeuron$Subcelltype <- factor(
  combined_seurat_ExcNeuron$Subcelltype,
  levels = paste0("Excit_", 1:4)
)

combined_seurat_Immune$Subcelltype <- factor(
  combined_seurat_Immune$Subcelltype,
  levels = paste0("Immune_", 1:3)
)
```

## UMAP plotting function
```{r}
make_subtype_umap <- function(
    object,
    colors,
    split_by_group = FALSE,
    coordinate_limits = NULL
) {
  plot <- DimPlot(
    object,
    reduction = "umap",
    group.by = "Subcelltype",
    split.by = if (split_by_group) "major_group" else NULL,
    label = TRUE,
    label.size = 6,
    repel = TRUE,
    pt.size = 0.5
  ) +
    scale_color_manual(
      values = colors,
      drop = FALSE
    ) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 5)
      )
    ) +
    theme_classic(base_family = "Arial") +
    theme(
      axis.text = element_text(size = 16, color = "black"),
      axis.title = element_text(size = 16, color = "black"),
      legend.position = "right",
      legend.text = element_text(size = 12, color = "black"),
      legend.title = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.5),
      axis.ticks.length = grid::unit(5, "pt")
    )

  if (!is.null(coordinate_limits)) {
    plot <- plot + coord_cartesian(
      xlim = coordinate_limits$x,
      ylim = coordinate_limits$y
    )
  }

  plot
}
```

## Generate and save UMAPs
```{r}
subtype_colors <- list(
  Astrocyte = c(
    "Astro_1" = "#FF5A5A",
    "Astro_2" = "#FF7794",
    "Astro_3" = "#FF8FC2",
    "Astro_4" = "#FFB0D8",
    "Astro_5" = "#FFD4C2"
  ),
  Excitatory = c(
    "Excit_1" = "#1F78FF",
    "Excit_2" = "#5CB6FF",
    "Excit_3" = "#9BD3FF",
    "Excit_4" = "#CFEAFF"
  ),
  Immune = c(
    "Immune_1" = "#FF7A3D",
    "Immune_2" = "#FFAA33",
    "Immune_3" = "#FFE066"
  )
)

reclustered_objects <- list(
  Astrocyte = combined_seurat_Astro,
  Excitatory = combined_seurat_ExcNeuron,
  Immune = combined_seurat_Immune
)

coordinate_limits <- list(
  Astrocyte = NULL,
  Excitatory = NULL,
  Immune = list(
    x = c(-10, 15),
    y = c(-8, 8)
  )
)

file_labels <- c(
  Astrocyte = "Astrocyte",
  Excitatory = "ExcNeuron",
  Immune = "Immune"
)

for (celltype in names(reclustered_objects)) {
  object <- reclustered_objects[[celltype]]

  umap_plot <- make_subtype_umap(
    object = object,
    colors = subtype_colors[[celltype]],
    coordinate_limits = coordinate_limits[[celltype]]
  )

  split_plot <- make_subtype_umap(
    object = object,
    colors = subtype_colors[[celltype]],
    split_by_group = TRUE,
    coordinate_limits = coordinate_limits[[celltype]]
  )

  print(umap_plot)
  print(split_plot)

  ggsave(
    filename = file.path(
      output_folder,
      paste0("UMAP_", file_labels[celltype], ".png")
    ),
    plot = umap_plot,
    width = 5,
    height = 4,
    units = "in",
    dpi = 300,
    bg = "white"
  )
}
```

## Marker-expression dot plot function
```{r}
make_dotplot <- function(
    seurat_object,
    selected_markers,
    marker_labels,
    high_color,
    show_y_axis = TRUE,
    show_legend = TRUE,
    panel_title = NULL
) {
  expression_data <- FetchData(
    seurat_object,
    vars = c(
      selected_markers,
      "Subcelltype",
      "major_group"
    ),
    layer = "counts"
  )

  subtype_levels <- levels(
    droplevels(expression_data$Subcelltype)
  )

  expression_long <- expression_data %>%
    pivot_longer(
      cols = all_of(selected_markers),
      names_to = "marker",
      values_to = "expression"
    )

  overall_marker_means <- expression_long %>%
    group_by(marker) %>%
    summarise(
      overall_marker_mean = mean(expression, na.rm = TRUE),
      .groups = "drop"
    )

  dotplot_data <- expression_long %>%
    group_by(marker, Subcelltype, major_group) %>%
    summarise(
      mean_expression = mean(expression, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(
      overall_marker_means,
      by = "marker"
    ) %>%
    group_by(marker) %>%
    mutate(
      relative_expression = ifelse(
        overall_marker_mean > 0,
        mean_expression / overall_marker_mean * 100,
        0
      ),
      scaled_expression = if (sd(mean_expression, na.rm = TRUE) > 0) {
        as.numeric(scale(mean_expression))
      } else {
        0
      }
    ) %>%
    ungroup() %>%
    mutate(
      marker = factor(
        marker,
        levels = rev(selected_markers)
      ),
      Subcelltype = factor(
        Subcelltype,
        levels = subtype_levels
      ),
      major_group = factor(
        major_group,
        levels = c("ALS", "Control")
      )
    )

  plot <- ggplot(
    dotplot_data,
    aes(
      x = Subcelltype,
      y = marker,
      size = relative_expression,
      fill = scaled_expression
    )
  ) +
    geom_point(
      shape = 21,
      color = "gray50",
      stroke = 0.15,
      alpha = 0.9
    ) +
    facet_grid(
      . ~ major_group,
      scales = "free_x",
      space = "free_x"
    ) +
    scale_y_discrete(
      limits = rev(selected_markers),
      labels = marker_labels
    ) +
    scale_x_discrete(
      limits = subtype_levels,
      drop = FALSE
    ) +
    scale_size_area(
      max_size = 8,
      limits = c(0, 600),
      breaks = c(75, 100, 200),
      labels = c("75%", "100%", "200%")
    ) +
    scale_fill_gradient2(
      low = "lightgray",
      mid = "white",
      high = high_color,
      limits = c(-1, 1),
      oob = scales::squish
    ) +
    labs(
      title = panel_title,
      size = "Relative\nexpression",
      fill = "Scaled\nexpression"
    ) +
    theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(size = 8, hjust = 0.5),
      axis.title = element_blank(),
      axis.text.x = element_text(
        size = 8,
        angle = 45,
        hjust = 1,
        vjust = 1,
        color = "black"
      ),
      axis.text.y = element_text(size = 8, color = "black"),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 8),
      strip.text = element_text(size = 8),
      strip.background = element_rect(fill = "grey90", color = NA),
      axis.line = element_line(color = "black", linewidth = 0.3),
      axis.ticks = element_line(color = "black", linewidth = 0.3),
      axis.ticks.length = grid::unit(3, "pt"),
      panel.spacing = grid::unit(0.3, "lines"),
      plot.margin = margin(5, 5, 5, 5)
    )

  if (!show_y_axis) {
    plot <- plot + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
  }

  if (!show_legend) {
    plot <- plot + theme(
      legend.position = "none"
    )
  }

  list(
    plot = plot,
    data = dotplot_data
  )
}
```

## Generate the combined marker-expression dot plot
```{r}
dotplot_settings <- list(
  Astrocyte = list(
    object = combined_seurat_Astro,
    color = "red",
    title = "Astrocyte"
  ),
  Immune = list(
    object = combined_seurat_Immune,
    color = "#FFAA00",
    title = "Immune"
  ),
  Excitatory = list(
    object = combined_seurat_ExcNeuron,
    color = "#005CE6",
    title = "Excitatory neuron"
  )
)

dotplot_results <- lapply(
  seq_along(dotplot_settings),
  function(i) {
    setting <- dotplot_settings[[i]]

    make_dotplot(
      seurat_object = setting$object,
      selected_markers = selected_markers,
      marker_labels = marker_labels,
      high_color = setting$color,
      show_y_axis = i == 1,
      show_legend = i == length(dotplot_settings),
      panel_title = setting$title
    )
  }
)

dotplots <- lapply(
  dotplot_results,
  `[[`,
  "plot"
)

n_subtypes <- vapply(
  dotplot_settings,
  function(setting) {
    nlevels(
      droplevels(setting$object$Subcelltype)
    )
  },
  integer(1)
)

combined_plot <- wrap_plots(
  dotplots,
  nrow = 1
) +
  plot_layout(widths = n_subtypes)

print(combined_plot)

ggsave(
  filename = file.path(
    output_folder,
    "MarkerExpression_Dotplot_Combined.png"
  ),
  plot = combined_plot,
  width = total_width,
  height = 2.5,
  units = "in",
  dpi = 300,
  limitsize = FALSE,
  bg = "white"
)
```

## Calculate the marker-expression score
```{r}
score_markers <- c(
  "Cell-154Sm-pTDP43",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-147Sm-Glutaminase",
  "Cell-159Tb-IL6ST"
)

calculate_score_data <- function(
    seurat_object,
    selected_markers
) {
  missing_markers <- setdiff(
    selected_markers,
    rownames(seurat_object[["Protein"]])
  )

  if (length(missing_markers) > 0) {
    stop(
      "Markers not found in the Protein assay: ",
      paste(missing_markers, collapse = ", ")
    )
  }

  expression_matrix <- as.matrix(
    GetAssayData(
      seurat_object,
      assay = "Protein",
      layer = "counts"
    )[selected_markers, , drop = FALSE]
  )

  expression_matrix <- asinh(expression_matrix)

  marker_means <- rowMeans(
    expression_matrix,
    na.rm = TRUE
  )

  marker_sds <- apply(
    expression_matrix,
    1,
    sd,
    na.rm = TRUE
  )

  marker_sds[is.na(marker_sds) | marker_sds == 0] <- 1

  scaled_matrix <- sweep(
    expression_matrix,
    1,
    marker_means,
    FUN = "-"
  )

  scaled_matrix <- sweep(
    scaled_matrix,
    1,
    marker_sds,
    FUN = "/"
  )

  shifted_matrix <- scaled_matrix + 3
  shifted_matrix[shifted_matrix < 0] <- 0

  score_data <- data.frame(
    Cell = colnames(shifted_matrix),
    score = colSums(shifted_matrix^2, na.rm = TRUE),
    Subcelltype = as.character(seurat_object$Subcelltype),
    major_group = as.character(seurat_object$major_group),
    stringsAsFactors = FALSE
  )

  subtype_levels <- levels(
    droplevels(seurat_object$Subcelltype)
  )

  score_data$Subcelltype <- factor(
    score_data$Subcelltype,
    levels = subtype_levels
  )

  score_data$major_group <- factor(
    score_data$major_group,
    levels = c("ALS", "Control")
  )

  score_data$group <- factor(
    paste(
      score_data$major_group,
      score_data$Subcelltype,
      sep = "."
    ),
    levels = c(
      paste("ALS", subtype_levels, sep = "."),
      paste("Control", subtype_levels, sep = ".")
    )
  )

  score_data <- score_data[
    !is.na(score_data$group),
    ,
    drop = FALSE
  ]

  score_summary <- score_data %>%
    group_by(group) %>%
    summarise(
      n_cells = n(),
      mean_score = mean(score, na.rm = TRUE),
      se_score = sd(score, na.rm = TRUE) / sqrt(n_cells),
      .groups = "drop"
    )

  list(
    cell_data = score_data,
    summary = score_summary,
    n_groups = nlevels(droplevels(score_data$group))
  )
}
```

## Score-plot function
```{r}
make_score_plots <- function(
    score_result,
    high_color,
    panel_title,
    show_y_axis = TRUE,
    show_legend = FALSE
) {
  score_data <- score_result$cell_data
  score_summary <- score_result$summary

  bar_plot <- ggplot(
    score_data,
    aes(
      x = group,
      y = score,
      color = major_group
    )
  ) +
    geom_jitter(
      alpha = 0.8,
      size = 0.1,
      width = 0.2
    ) +
    geom_crossbar(
      data = score_summary,
      aes(
        x = group,
        y = mean_score,
        ymin = mean_score,
        ymax = mean_score
      ),
      inherit.aes = FALSE,
      width = 0.6,
      linewidth = 0.3,
      color = "black"
    ) +
    geom_errorbar(
      data = score_summary,
      aes(
        x = group,
        ymin = mean_score - se_score,
        ymax = mean_score + se_score
      ),
      inherit.aes = FALSE,
      width = 0.4,
      linewidth = 0.6,
      color = "black"
    ) +
    scale_color_manual(
      values = c(
        "ALS" = "red",
        "Control" = "blue"
      )
    ) +
    scale_y_continuous(
      expand = c(0, 0)
    ) +
    labs(
      title = panel_title,
      x = NULL,
      y = "Expression score (AU)"
    ) +
    theme_classic(base_size = 16) +
    theme(
      plot.title = element_text(
        size = 12,
        color = "black",
        face = "bold",
        hjust = 0.5
      ),
      axis.text.x = element_text(
        size = 8,
        color = "black",
        angle = 45,
        hjust = 1
      ),
      axis.text.y = element_text(size = 10, color = "black"),
      axis.title.y = element_text(size = 14, color = "black"),
      legend.position = "none",
      axis.line = element_line(color = "black", linewidth = 0.5),
      axis.ticks.length = grid::unit(5, "pt")
    )

  if (!show_y_axis) {
    bar_plot <- bar_plot + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank()
    )
  }

  heatmap_plot <- ggplot(
    score_summary,
    aes(
      x = group,
      y = 1,
      fill = mean_score
    )
  ) +
    geom_tile(color = "white") +
    scale_fill_gradient(
      low = "white",
      high = high_color,
      name = "Expression\nscore"
    ) +
    scale_y_continuous(
      expand = c(0, 0)
    ) +
    labs(
      title = panel_title,
      x = NULL,
      y = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(
        size = 12,
        color = "black",
        hjust = 0.5
      ),
      axis.text.x = element_text(
        angle = 45,
        hjust = 1,
        size = 8,
        color = "black"
      ),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid = element_blank(),
      legend.position = if (show_legend) "right" else "none",
      legend.title = element_text(size = 9, color = "black"),
      legend.text = element_text(size = 8, color = "black")
    )

  list(
    bar_plot = bar_plot,
    heatmap_plot = heatmap_plot
  )
}
```

## Generate combined score plots
```{r}
score_settings <- list(
  Astrocyte = list(
    object = combined_seurat_Astro,
    color = "red",
    title = "Astrocyte"
  ),
  Immune = list(
    object = combined_seurat_Immune,
    color = "#FFAA00",
    title = "Immune"
  ),
  Excitatory = list(
    object = combined_seurat_ExcNeuron,
    color = "#005CE6",
    title = "Excitatory neuron"
  )
)

score_results <- lapply(
  score_settings,
  function(setting) {
    calculate_score_data(
      seurat_object = setting$object,
      selected_markers = score_markers
    )
  }
)

score_plots <- lapply(
  seq_along(score_settings),
  function(i) {
    make_score_plots(
      score_result = score_results[[i]],
      high_color = score_settings[[i]]$color,
      panel_title = score_settings[[i]]$title,
      show_y_axis = i == 1,
      show_legend = i == length(score_settings)
    )
  }
)

n_groups <- vapply(
  score_results,
  `[[`,
  integer(1),
  "n_groups"
)

bar_combined <- wrap_plots(
  lapply(score_plots, `[[`, "bar_plot"),
  nrow = 1
) +
  plot_layout(widths = n_groups)

heatmap_combined <- wrap_plots(
  lapply(score_plots, `[[`, "heatmap_plot"),
  nrow = 1
) +
  plot_layout(widths = n_groups)

print(bar_combined)
print(heatmap_combined)

ggsave(
  filename = file.path(
    output_folder,
    "Expression_score_heatmap_Combined.png"
  ),
  plot = heatmap_combined,
  width = total_width,
  height = 1.6,
  units = "in",
  dpi = 300,
  limitsize = FALSE,
  bg = "white"
)
```