# =============================================================================
# 5_IMC-Hyperion_Monocle_trajectory.Rmd
#
# Description:
# This script infers and visualizes a predicted cellular continuum using
# Monocle3 pseudotime analysis.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Monocle3 cell trajectory object
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(ggplot2)
  library(monocle3)
  library(SeuratWrappers)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load the integrated Seurat object and select ALS astrocytes
```{r}
combined_seurat<-readRDS("combined_seurat_PrecentralGyrus.rds")

astro_als <- subset(
  combined_seurat,
  subset = Celltype == "Astrocyte" & major_group == "ALS"
)
```

## Calculate the marker-based score used to select the root cell
```{r}
root_markers <- c(
  "Cell-154Sm-pTDP43",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-147Sm-Glutaminase",
  "Cell-159Tb-IL6ST"
)

protein_counts <- GetAssayData(
  astro_als,
  assay = "Protein",
  layer = "counts"
)

missing_markers <- setdiff(root_markers, rownames(protein_counts))
if (length(missing_markers) > 0) {
  stop("Missing root-score markers: ", paste(missing_markers, collapse = ", "))
}

marker_matrix <- asinh(
  as.matrix(protein_counts[root_markers, , drop = FALSE])
)

# Z-score each marker across cells. Markers with zero variance are assigned 0.
marker_z <- t(scale(t(marker_matrix)))
marker_z[!is.finite(marker_z)] <- 0

marker_shifted <- marker_z + 3
marker_shifted[marker_shifted < 0] <- 0

root_score <- colSums(marker_shifted^2)
root_cell <- names(which.min(root_score))
```

## Construct the Monocle3 trajectory
```{r}
cds <- as.cell_data_set(astro_als)

reducedDims(cds)$PCA <- Embeddings(astro_als, reduction = "harmony")

colData(cds)$Subcelltype <- astro_als$Subcelltype
colData(cds)$major_group <- astro_als$major_group

cds <- reduce_dimension(
  cds,
  reduction_method = "UMAP",
  preprocess_method = "PCA",
  umap.min_dist = 0.001,
  umap.n_neighbors = 10,
  umap.nn_method = "annoy",
  umap.fast_sgd = FALSE,
  cores = 1
)

cds <- cluster_cells(cds, reduction_method = "UMAP")

cds <- learn_graph(
  cds,
  use_partition = TRUE,
  learn_graph_control = list(minimal_branch_len = 15)
)

cds <- order_cells(cds, root_cells = root_cell)
```

## Plotting settings
```{r}
subcelltype_colors <- c(
  "Astro_1" = "#FF5A5A",
  "Astro_2" = "#FF7794",
  "Astro_3" = "#FF8FC2",
  "Astro_4" = "#FFB0D8",
  "Astro_5" = "#FFD4C2"
)

trajectory_theme <- theme_classic() +
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.title = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", linewidth = 0.5),
    axis.ticks.length = grid::unit(5, "pt"),
    strip.text = element_text(size = 20, color = "black")
  )

save_trajectory_plot <- function(plot_object, file_stem) {
  ggsave(
    filename = file.path(output_folder, paste0(file_stem, ".png")),
    plot = plot_object,
    width = 8,
    height = 8,
    units = "in",
    dpi = 300,
    limitsize = FALSE
  )
}
```

## Pseudotime trajectory
```{r}
p_pseudotime <- plot_cells(
  cds,
  color_cells_by = "pseudotime",
  label_cell_groups = FALSE,
  group_cells_by = "Subcelltype",
  label_leaves = FALSE,
  label_branch_points = FALSE,
  label_roots = FALSE,
  cell_size = 2.5,
  show_trajectory_graph = TRUE,
  trajectory_graph_color = "green",
  trajectory_graph_segment_size = 2
) +
  coord_cartesian(xlim = c(-6, 6), ylim = c(-6, 6)) +
  scale_x_reverse() +
  trajectory_theme

print(p_pseudotime)
save_trajectory_plot(p_pseudotime, "Celltransition_Astrocyte_1")
```

## Subcell-type trajectory
```{r}
observed_subcelltypes <- unique(as.character(colData(cds)$Subcelltype))
missing_colors <- setdiff(observed_subcelltypes, names(subcelltype_colors))

if (length(missing_colors) > 0) {
  stop(
    "No plotting colors were specified for: ",
    paste(missing_colors, collapse = ", ")
  )
}

p_subcelltype <- plot_cells(
  cds,
  color_cells_by = "Subcelltype",
  label_cell_groups = FALSE,
  group_cells_by = "Subcelltype",
  label_leaves = FALSE,
  label_branch_points = FALSE,
  label_roots = FALSE,
  cell_size = 2.5,
  show_trajectory_graph = TRUE,
  trajectory_graph_color = "green",
  trajectory_graph_segment_size = 2
) +
  scale_color_manual(
    values = subcelltype_colors,
    name = "Subcelltype"
  ) +
  coord_cartesian(xlim = c(-6, 6), ylim = c(-6, 6)) +
  scale_x_reverse() +
  trajectory_theme

print(p_subcelltype)
save_trajectory_plot(p_subcelltype, "Celltransition_Astrocyte_2")
```

