# =============================================================================
# IMC_marker_expression_profiles.R
#
# Description:
# This script generates dot plots showing marker expression across
# subcelltypes for selected major cell classes.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Dot plots summarizing marker expression (pTDP43, UBQLN2, GLS, TREM2)
#   across subcelltypes
# - Heatmap visualization of marker expression patterns
#
# Notes:
# Analysis parameters are currently configured for the Precentral gyrus dataset.
# For analyses of other brain regions, please refer to "IMC_analysis_paramaters.txt"
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================


## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
  library(tidyr)
  library(scales)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load previously saved object
```{r}
combined_seurat <- readRDS(
   file = file.path("rds/combined_seurat_PrecentralGyrus.rds")
)
```

## Markers info
```{r}
Markers_astro <- c("Cell-141Pr-AQP4", "Cell-158Gd-GFAP")
Markers_immune <- c("Cell-161Dy-CX3CR1", "Cell-164Dy-TMEM119","Cell-160Gd-IBA1","Cell-163Dy-CD4", "Cell-155Gd-FoxP3", "Cell-167Er-CD8")
Markers_neuron <- c("Cell-151Eu-MAP2","Cell-149Sm-C9orf72", "Cell-169Tm-ChAT","Cell-166Er-GAD67")
Markers_others <- c("Cell-175Lu-TDP43",
"Cell-154Sm-pTDP43",
"Cell-168Er-FUS",
"Cell-152Sm-ATXN2",
"Cell-145Nd-UBQLN2",
"Cell-170Er-APOE",
"Cell-176Yb-TBK1",
"Cell-162Dy-pTBK1",
"Cell-147Sm-Glutaminase",
"Cell-143Nd-SQSTM1",
"Cell-156Gd-CH24H",
"Cell-153Eu-ORF1",
"Cell-159Tb-IL6ST",
"Cell-165Ho-PD1",
"Cell-150Nd-PDL1",
"Cell-172Yb-PDL2",
"Cell-171Yb-TREM2",
"Cell-174Yb-MHCclassII"
)

Markers <- c(Markers_astro, Markers_immune, Markers_neuron, Markers_others)

Marker_label <- c(
  "Cell-141Pr-AQP4" = "AQP4",
  "Cell-143Nd-SQSTM1" = "SQSTM1",
  "Cell-145Nd-UBQLN2" = "UBQLN2",
  "Cell-147Sm-Glutaminase" = "GLS",
  "Cell-149Sm-C9orf72" = "NeuN",
  "Cell-150Nd-PDL1" = "PDL1",
  "Cell-151Eu-MAP2" = "MAP2",
  "Cell-152Sm-ATXN2" = "ATXN2",
  "Cell-153Eu-ORF1" = "ORF1",
  "Cell-154Sm-pTDP43" = "pTDP43",
  "Cell-155Gd-FoxP3" = "FOXP3",
  "Cell-156Gd-CH24H" = "CH24H",
  "Cell-158Gd-GFAP" = "GFAP",
  "Cell-159Tb-IL6ST" = "IL6ST",
  "Cell-160Gd-IBA1" = "IBA1",
  "Cell-161Dy-CX3CR1" = "CX3CR1",
  "Cell-162Dy-pTBK1" = "pTBK1",
  "Cell-163Dy-CD4" = "CD4",
  "Cell-164Dy-TMEM119" = "TMEM119",
  "Cell-165Ho-PD1" = "PD1",
  "Cell-166Er-GAD67" = "GAD67",
  "Cell-167Er-CD8" = "CD8",
  "Cell-168Er-FUS" = "FUS",
  "Cell-169Tm-ChAT" = "ChAT",
  "Cell-170Er-APOE" = "APOE",
  "Cell-171Yb-TREM2" = "TREM2",
  "Cell-172Yb-PDL2" = "PDL2",
  "Cell-174Yb-MHCclassII" = "MHCII",
  "Cell-175Lu-TDP43" = "TDP43",
  "Cell-176Yb-TBK1" = "TBK1"
)

```

## Dotplot
# Astro
```{r}
sel_markers <- c("Cell-154Sm-pTDP43",
                 "Cell-145Nd-UBQLN2",
                 "Cell-147Sm-Glutaminase",
                 "Cell-171Yb-TREM2")

marker_labels <- c(
  "Cell-154Sm-pTDP43"      = "pTDP43",
  "Cell-145Nd-UBQLN2"      = "UBQLN2",
  "Cell-147Sm-Glutaminase" = "GLS",
  "Cell-171Yb-TREM2"       = "TREM2"
)

temp <- subset(combined_seurat, subset = Celltype %in% c("Astrocyte"))

vln_data <- FetchData(
  temp,
  vars = c(sel_markers, "Subcelltype", "major_group"),
  slot = "data"
)


sub_levels <- vln_data %>%
  distinct(Subcelltype) %>%
  pull(Subcelltype) %>%
  unique()

sub_levels <- c("Astro_1","Astro_2","Astro_3","Astro_4","Astro_5", "") 

if (all(grepl("^[A-Za-z]+\\d+$", sub_levels))) {
  sub_levels <- sub_levels[order(as.integer(gsub("^[A-Za-z]+", "", sub_levels)))]
} else {
  sub_levels <- sort(sub_levels)
}


mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker, Subcelltype, major_group) %>%
  summarise(mean_expr = mean(expr, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    marker     = factor(marker, levels = rev(sel_markers)),
    Subcelltype = factor(Subcelltype, levels = sub_levels),
    major_group = factor(major_group, levels = c("ALS", "Control"))
  )

marker_mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker) %>%
  summarise(marker_mean_expr = mean(expr, na.rm = TRUE),
            .groups = "drop")

mean_df <- mean_df %>%
  left_join(marker_mean_df, by = "marker") %>%
  mutate(
    percent_expr = (mean_expr / marker_mean_expr) * 100
  )

dot_df_scaled <- mean_df %>%
  group_by(marker) %>%
  mutate(
    mean_expr_scaled = scale(mean_expr)[,1],
     ) %>%
  ungroup()

dot_df_scaled <- dot_df_scaled %>%
  mutate(Subcelltype = factor(Subcelltype, levels = sub_levels))

p <- ggplot(dot_df_scaled,
            aes(x = Subcelltype,
                y = marker,
                size = percent_expr,
                color = mean_expr_scaled)) +
  geom_point(alpha = 0.9) +
  facet_grid(. ~ major_group, scales = "free_x", space = "free_x") +
 scale_y_discrete(
  limits = rev(sel_markers),
  labels = marker_labels
)+
  scale_x_discrete(limits = sub_levels, drop = FALSE) +
 scale_size_area(
  max_size = 12,
     limits = c(0, 200),
    breaks = c(75, 100, 200),
  labels = c("75%", "100%", "200%"),
    oob = scales::squish
  ) +  
  scale_color_gradient2(low = "lightgray", high = "red", mid = "white", 
                       limits = c(-1.5, 1.5), oob = scales::squish) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.x = element_blank(),# axis.title.y = element_text(size = 10),
     axis.title.y = element_blank(), #axis.title.x = element_text(size = 10),
     axis.text.x = element_blank(), #axis.text.x  = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_blank(), #axis.text.y  = element_text(size = 10),
    legend.title = element_text(size = 6),
    legend.text  = element_text(size = 6),
    plot.title   = element_text(size = 12),
    axis.line = element_line(color = "black", linewidth = 0.5)
  ) +
  labs(size="Mean expression\n(%)", color="Z-scaled\nexpression")

print(p)

```

# Immune
```{r}
temp <- subset(combined_seurat, subset = Celltype %in% c("Immune"))

vln_data <- FetchData(
  temp,
  vars = c(sel_markers, "Subcelltype", "major_group"),
  slot = "data"
)


sub_levels <- vln_data %>%
  distinct(Subcelltype) %>%
  pull(Subcelltype) %>%
  unique()

sub_levels <- c("Immune_1","Immune_2","Immune_3",""," ", "   ") 

if (all(grepl("^[A-Za-z]+\\d+$", sub_levels))) {
  sub_levels <- sub_levels[order(as.integer(gsub("^[A-Za-z]+", "", sub_levels)))]
} else {
  sub_levels <- sort(sub_levels)
}


mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker, Subcelltype, major_group) %>%
  summarise(mean_expr = mean(expr, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    marker     = factor(marker, levels = rev(sel_markers)),
    Subcelltype = factor(Subcelltype, levels = sub_levels),
    major_group = factor(major_group, levels = c("ALS", "Control"))
  )

marker_mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker) %>%
  summarise(marker_mean_expr = mean(expr, na.rm = TRUE),
            .groups = "drop")

mean_df <- mean_df %>%
  left_join(marker_mean_df, by = "marker") %>%
  mutate(
    percent_expr = (mean_expr / marker_mean_expr) * 100
  )

dot_df_scaled <- mean_df %>%
  group_by(marker) %>%
  mutate(
    mean_expr_scaled = scale(mean_expr)[,1],
     ) %>%
  ungroup()

dot_df_scaled <- dot_df_scaled %>%
  mutate(Subcelltype = factor(Subcelltype, levels = sub_levels))

p <- ggplot(dot_df_scaled,
            aes(x = Subcelltype,
                y = marker,
                size = percent_expr,
                color = mean_expr_scaled)) +
  geom_point(alpha = 0.9) +
  facet_grid(. ~ major_group, scales = "free_x", space = "free_x") +
 scale_y_discrete(
  limits = rev(sel_markers),
  labels = marker_labels
)+
  scale_x_discrete(limits = sub_levels, drop = FALSE) +
 scale_size_area(
  max_size = 12,
     limits = c(0, 200),
    breaks = c(75, 100, 200),
  labels = c("75%", "100%", "200%"),
    oob = scales::squish
  ) +  
  scale_color_gradient2(low = "lightgray", high = "#FFAA00",  mid = "white", 
                       limits = c(-1.5, 1.5), oob = scales::squish) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.x = element_blank(),# axis.title.y = element_text(size = 10),
     axis.title.y = element_blank(), #axis.title.x = element_text(size = 10),
     axis.text.x = element_blank(), #axis.text.x  = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_blank(), #axis.text.y  = element_text(size = 10),
    legend.title = element_text(size = 6),
    legend.text  = element_text(size = 6),
    plot.title   = element_text(size = 12),
    axis.line = element_line(color = "black", linewidth = 0.5)
  ) +
  labs(size="Mean expression\n(%)", color="Z-scaled\nexpression")

print(p)
```

# Excitatory
```{r}

temp <- subset(combined_seurat, subset = Celltype %in% c("Excitatory"))

vln_data <- FetchData(
  temp,
  vars = c(sel_markers, "Subcelltype", "major_group"),
  slot = "data"
)


sub_levels <- vln_data %>%
  distinct(Subcelltype) %>%
  pull(Subcelltype) %>%
  unique()

sub_levels <- c("Exc_1","Exc_2","Exc_3","Exc_4","Exc_5","") 

if (all(grepl("^[A-Za-z]+\\d+$", sub_levels))) {
  sub_levels <- sub_levels[order(as.integer(gsub("^[A-Za-z]+", "", sub_levels)))]
} else {
  sub_levels <- sort(sub_levels)
}


mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker, Subcelltype, major_group) %>%
  summarise(mean_expr = mean(expr, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    marker     = factor(marker, levels = rev(sel_markers)),
    Subcelltype = factor(Subcelltype, levels = sub_levels),
    major_group = factor(major_group, levels = c("ALS", "Control"))
  )

marker_mean_df <- vln_data %>%
  pivot_longer(cols = all_of(sel_markers),
               names_to = "marker",
               values_to = "expr") %>%
  group_by(marker) %>%
  summarise(marker_mean_expr = mean(expr, na.rm = TRUE),
            .groups = "drop")

mean_df <- mean_df %>%
  left_join(marker_mean_df, by = "marker") %>%
  mutate(
    percent_expr = (mean_expr / marker_mean_expr) * 100
  )

dot_df_scaled <- mean_df %>%
  group_by(marker) %>%
  mutate(
    mean_expr_scaled = scale(mean_expr)[,1],
     ) %>%
  ungroup()

dot_df_scaled <- dot_df_scaled %>%
  mutate(Subcelltype = factor(Subcelltype, levels = sub_levels))

p <- ggplot(dot_df_scaled,
            aes(x = Subcelltype,
                y = marker,
                size = percent_expr,
                color = mean_expr_scaled)) +
  geom_point(alpha = 0.9) +
  facet_grid(. ~ major_group, scales = "free_x", space = "free_x") +
 scale_y_discrete(
  limits = rev(sel_markers),
  labels = marker_labels
)+
  scale_x_discrete(limits = sub_levels, drop = FALSE) +
 scale_size_area(
  max_size = 12,
     limits = c(0, 200),
    breaks = c(75, 100, 200),
  labels = c("75%", "100%", "200%"),
    oob = scales::squish
  ) +  
  scale_color_gradient2(low = "lightgray", high = "#005CE6",  mid = "white", 
                       limits = c(-1.5, 1.5), oob = scales::squish) +
  theme_classic(base_size = 14) +
  theme(
    axis.title.x = element_blank(),# axis.title.y = element_text(size = 10),
     axis.title.y = element_blank(), #axis.title.x = element_text(size = 10),
     axis.text.x = element_blank(), #axis.text.x  = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_blank(), #axis.text.y  = element_text(size = 10),
    legend.title = element_text(size = 6),
    legend.text  = element_text(size = 6),
    plot.title   = element_text(size = 12),
    axis.line = element_line(color = "black", linewidth = 0.5)
  ) +
  labs(size="Mean expression\n(%)", color="Z-scaled\nexpression")

print(p)

```

## Euclidean distance
# Astro
```{r}

sel_markers <- c("Cell-154Sm-pTDP43",
                 "Cell-147Sm-Glutaminase",
                 "Cell-145Nd-UBQLN2",
                 "Cell-171Yb-TREM2")

temp <- subset(combined_seurat, subset = Celltype %in% c("Astrocyte"))
mat <- GetAssayData(temp, assay = "Protein", slot = "counts")[sel_markers, ]


mat_z <- t(scale(t(mat))) 
df <- as.data.frame(t(mat_z)) 
mat_shift <- mat_z + 3

mat_shift[mat_shift < 0] <- 0
df <- as.data.frame(t(mat_shift))


for (marker in sel_markers) {
  hist(
    mat_z[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored", marker),
    xlab = "Z-score",
    border = "white"
  )
}

for (marker in sel_markers) {
  hist(
    mat_shift[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored+3", marker),
    xlab = "Z-score",
    border = "white"
  )
}



df$score <- apply(df, 1, function(x) sum(x^2))

df$Subcelltype <- temp$Subcelltype
df$major_group <- temp$major_group
df$group<-interaction(df$major_group, df$Subcelltype, drop = TRUE)



df$group <- factor(
  df$group,
  levels = c(
     paste0("ALS.Astro_", 1:5),
      paste0("Control.Astro_", 1:5)))


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop")


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(
    mean_score = mean(score, na.rm = TRUE),
    se_score   = sd(score, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups = "drop"
  )
   

dodge_width <- 0.8

p <- ggplot(df,
            aes(x = group,
                y = score,
                color = major_group)) +
  geom_jitter(
    alpha = 0.8,
    size  = 0.1,
    width = 0.2
  ) +
  geom_crossbar(
    data = mean_score_df,
    aes(x = group,
        y = mean_score,
        ymin = mean_score,
        ymax = mean_score),
    inherit.aes = FALSE,
    width = 0.6,
    linewidth = 0.3,
    color = "black"
  ) +
  geom_errorbar(
  data = mean_score_df,
  aes(
    x = group,
    ymin = mean_score - se_score,
    ymax = mean_score + se_score
  ),
   inherit.aes = FALSE,
  width = 0.4,
  color = "black",
  linewidth = 0.6
)+
  scale_color_manual(values = c("ALS" = "red", "Control" = "blue")) +
scale_y_continuous(expand = c(0,0))+
  theme_classic(base_size = 16) +
  labs(
    title = "Expression (sum x^2)",
    x= NULL,
    y = "AU"
  ) +
  theme(
    plot.title      = element_text(size = 12, color = "black"),
    axis.text.x     = element_text(size = 8, color = "black", angle=45, hjust=1),
    axis.text.y     = element_text(size = 10, color = "black"),
    axis.title.y    = element_text(size = 14, color = "black"),
    legend.position = "none",
    axis.line       = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt")
  )

print(p)


df$group <- factor(df$group)
result <- aov(score ~ group, data = df)
summary(result)

TukeyHSD(result)



p<-ggplot(mean_score_df, aes(x = group, y = 1, fill = mean_score)) +
  geom_tile(color = "white") +
   scale_fill_gradient(
    low  = "white",
    high = "red",
      name = "Expression (sum x^2)"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.title.y  = element_blank(),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.title.x  = element_blank(),
    axis.text.x   = element_text(angle = 45, hjust = 1, size = 8),
    panel.grid    = element_blank()
  )

print(p)


 write.csv(
    df,
    file = paste0(output_folder, "Marker_expression_score_Astrocyte.csv"),
    row.names = FALSE
)
```

# Immune
```{r}
temp <- subset(combined_seurat, subset = Celltype %in% c("Immune"))
mat <- GetAssayData(temp, assay = "Protein", slot = "counts")[sel_markers, ]


mat_z <- t(scale(t(mat))) 
df <- as.data.frame(t(mat_z)) 
mat_shift <- mat_z + 3

mat_shift[mat_shift < 0] <- 0
df <- as.data.frame(t(mat_shift))


for (marker in sel_markers) {
  hist(
    mat_z[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored", marker),
    xlab = "Z-score",
    border = "white"
  )
}

for (marker in sel_markers) {
  hist(
    mat_shift[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored+3", marker),
    xlab = "Z-score",
    border = "white"
  )
}



df$score <- apply(df, 1, function(x) sum(x^2))

df$Subcelltype <- temp$Subcelltype
df$major_group <- temp$major_group
df$group<-interaction(df$major_group, df$Subcelltype, drop = TRUE)



df$group <- factor(
  df$group,
  levels = c(
     paste0("ALS.Immune_", 1:5),
      paste0("Control.Immune_", 1:5)))


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop")


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(
    mean_score = mean(score, na.rm = TRUE),
    se_score   = sd(score, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups = "drop"
  )
   

dodge_width <- 0.8

p <- ggplot(df,
            aes(x = group,
                y = score,
                color = major_group)) +
  geom_jitter(
    alpha = 0.8,
    size  = 0.1,
    width = 0.2
  ) +
  geom_crossbar(
    data = mean_score_df,
    aes(x = group,
        y = mean_score,
        ymin = mean_score,
        ymax = mean_score),
    inherit.aes = FALSE,
    width = 0.6,
    linewidth = 0.3,
    color = "black"
  ) +
  geom_errorbar(
  data = mean_score_df,
  aes(
    x = group,
    ymin = mean_score - se_score,
    ymax = mean_score + se_score
  ),
   inherit.aes = FALSE,
  width = 0.4,
  color = "black",
  linewidth = 0.6
)+
  scale_color_manual(values = c("ALS" = "red", "Control" = "blue")) +
scale_y_continuous(expand = c(0,0))+
  theme_classic(base_size = 16) +
  labs(
    title = "Expression (sum x^2)",
    x= NULL,
    y = "AU"
  ) +
  theme(
    plot.title      = element_text(size = 12, color = "black"),
    axis.text.x     = element_text(size = 8, color = "black", angle=45, hjust=1),
    axis.text.y     = element_text(size = 10, color = "black"),
    axis.title.y    = element_text(size = 14, color = "black"),
    legend.position = "none",
    axis.line       = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt")
  )

print(p)

 
df$group <- factor(df$group)
result <- aov(score ~ group, data = df)
summary(result)

TukeyHSD(result)

p<-ggplot(mean_score_df, aes(x = group, y = 1, fill = mean_score)) +
    geom_tile(color = "white") +
   scale_fill_gradient(
    low  = "white",
    high = "#FFAA00",
    name = "Expression (sum x^2)"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.title.y  = element_blank(),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.title.x  = element_blank(),
    axis.text.x   = element_text(angle = 45, hjust = 1, size = 8),
    panel.grid    = element_blank()
  )

print(p)


write.csv(
    df,
    file = paste0(output_folder, "Marker_expression_score_Immune.csv"),
    row.names = FALSE
)
```

# Excitatory
```{r}

temp <- subset(combined_seurat, subset = Celltype %in% c("Excitatory"))
mat <- GetAssayData(temp, assay = "Protein", slot = "counts")[sel_markers, ]


mat_z <- t(scale(t(mat))) 
df <- as.data.frame(t(mat_z)) 
mat_shift <- mat_z + 3

mat_shift[mat_shift < 0] <- 0
df <- as.data.frame(t(mat_shift))


for (marker in sel_markers) {
  hist(
    mat_z[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored", marker),
    xlab = "Z-score",
    border = "white"
  )
}

for (marker in sel_markers) {
  hist(
    mat_shift[marker, ], 
    breaks = 40,
    col = "orange",
    main = paste("Z-scored+3", marker),
    xlab = "Z-score",
    border = "white"
  )
}



df$score <- apply(df, 1, function(x) sum(x^2))

df$Subcelltype <- temp$Subcelltype
df$major_group <- temp$major_group
df$group<-interaction(df$major_group, df$Subcelltype, drop = TRUE)



df$group <- factor(
  df$group,
  levels = c(
     paste0("ALS.Exc_", 1:5),
      paste0("Control.Exc_", 1:5)))


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop")


mean_score_df <- df %>%
  group_by(group) %>%
  summarise(
    mean_score = mean(score, na.rm = TRUE),
    se_score   = sd(score, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups = "drop"
  )
   

dodge_width <- 0.8

p <- ggplot(df,
            aes(x = group,
                y = score,
                color = major_group)) +
  geom_jitter(
    alpha = 0.8,
    size  = 0.1,
    width = 0.2
  ) +
  geom_crossbar(
    data = mean_score_df,
    aes(x = group,
        y = mean_score,
        ymin = mean_score,
        ymax = mean_score),
    inherit.aes = FALSE,
    width = 0.6,
    linewidth = 0.3,
    color = "black"
  ) +
  geom_errorbar(
  data = mean_score_df,
  aes(
    x = group,
    ymin = mean_score - se_score,
    ymax = mean_score + se_score
  ),
   inherit.aes = FALSE,
  width = 0.4,
  color = "black",
  linewidth = 0.6
)+
  scale_color_manual(values = c("ALS" = "red", "Control" = "blue")) +
scale_y_continuous(expand = c(0,0))+
  theme_classic(base_size = 16) +
  labs(
    title = "Expression (sum x^2)",
    x= NULL,
    y = "AU"
  ) +
  theme(
    plot.title      = element_text(size = 12, color = "black"),
    axis.text.x     = element_text(size = 8, color = "black", angle=45, hjust=1),
    axis.text.y     = element_text(size = 10, color = "black"),
    axis.title.y    = element_text(size = 14, color = "black"),
    legend.position = "none",
    axis.line       = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt")
  )

print(p)


df$group <- factor(df$group)
result <- aov(score ~ group, data = df)
summary(result)

TukeyHSD(result)

p<-ggplot(mean_score_df, aes(x = group, y = 1, fill = mean_score)) +
  geom_tile(color = "white") +
   scale_fill_gradient(
    low  = "white",
    high = "#005CE6",
    name = "Expression (sum x^2)"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.title.y  = element_blank(),
    axis.text.y   = element_blank(),
    axis.ticks.y  = element_blank(),
    axis.title.x  = element_blank(),
    axis.text.x   = element_text(angle = 45, hjust = 1, size = 8),
    panel.grid    = element_blank()
  )

print(p)

write.csv(
    df,
    file = paste0(output_folder, "Marker_expression_score_Exc.csv"),
    row.names = FALSE
)
```
