# =============================================================================
# IMC_cell_transition.R
#
# Description:
# This script performs cell-state transition analysis of astrocyte subcelltypes.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Monocle3 cell trajectory object
#
# Notes:
# This script is currently configured for astrocytes from the Precentral gyrus
# dataset. Root cells are selected from the Astro_4 subcelltype based on the
# cell closest to the UMAP centroid of that subpopulation.
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# Redistribution should retain this header.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

# Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
  library(monocle3)
library(SeuratWrappers)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load previously saved object
```{r}
combined_seurat <- readRDS(
   file = file.path("rds/combined_seurat_PrecentralGyrus.rds")
)

combined_seurat_Astro <- subset(combined_seurat, subset = Celltype == "Astrocyte")
```


## Cell transition 
```{r}
combined_seurat_Astro <- subset(combined_seurat, subset = Celltype == "Astrocyte")
temp <- subset(combined_seurat_Astro, subset = major_group == "ALS")

cds <- as.cell_data_set(temp)

reducedDims(cds)$PCA <- Embeddings(temp, "harmony")

colData(cds)$Subcelltype <- temp$Subcelltype
colData(cds)$major_group <- temp$major_group

cds <- reduce_dimension(
  cds,
  reduction_method = "UMAP",
  preprocess_method = "PCA",
  umap.min_dist = 0.001,          
  umap.n_neighbors = 5,         
  umap.nn_method = "annoy",     
  umap.fast_sgd = FALSE         
)


cds <- cluster_cells(cds)
cds <- learn_graph(cds)

astro4_cells <- colnames(cds)[cds@colData$Subcelltype == "Astro_4"]


coords <- reducedDims(cds)$UMAP[astro4_cells, ]


center <- colMeans(coords)
dist_to_center <- apply(coords, 1, function(x) sum((x - center)^2))

root_id <- names(which.min(dist_to_center))

cds <- order_cells(cds, root_cells = root_id)



p <- plot_cells(
  cds,
  color_cells_by = "pseudotime",
   label_cell_groups = FALSE,
  label_leaves = FALSE,         
  label_branch_points = FALSE,  
  label_roots = FALSE,    
  graph_label_size = 3
)

plot(p)




p <- plot_cells(
  cds,
  color_cells_by = "Subcelltype",
  label_cell_groups = FALSE,
  label_leaves = FALSE,         
  label_branch_points = FALSE,  
  label_roots = FALSE           
) + 
  theme_classic() 

print(p)

p <- plot_cells(
  cds,
  color_cells_by = "pseudotime",
    label_cell_groups     = FALSE,         
  group_cells_by        = "Subcelltype", 
  label_leaves          = FALSE,         
  label_branch_points   = FALSE,  
  graph_label_size      = 0,             
  cell_size             = 1.5, 
  show_trajectory_graph = TRUE
) +
  coord_cartesian(xlim = c(-8, 8), ylim = c(-7, 7)) + 
  theme_classic() +
  theme(
    axis.text   = element_text(size = 20, color = "black"),
    axis.title  = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.title= element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line   = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text  = element_text(size = 20, color = "black")
  )

plot(p)

celltype_colors <- c(
  "Astro_1" = "#FF5A5A",   
  "Astro_2" = "#FF7794",  
  "Astro_3" = "#FF8FC2", 
  "Astro_4" = "#FFB0D8", 
  "Astro_5" = "#FFD4C2"
)

p <- plot_cells(
  cds,
  color_cells_by        = "Subcelltype",
  label_cell_groups     = FALSE,         
  group_cells_by        = "Subcelltype", 
  label_leaves          = FALSE,         
  label_branch_points   = FALSE,  
  graph_label_size      = 0,             
  cell_size             = 1.5,           
  show_trajectory_graph = TRUE
) +
  scale_color_manual(
    values = celltype_colors,
    name   = "Subcelltype"
  ) +
  coord_cartesian(xlim = c(-8, 8), ylim = c(-7, 7)) + 
  theme_classic() +
  theme(
    axis.text   = element_text(size = 20, color = "black"),
    axis.title  = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.title= element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line   = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text  = element_text(size = 20, color = "black")
  )

print(p)


```
