# =============================================================================
# IMC_DEPs_heatmap.R
#
# Description:
# This script computes differential marker expression between ALS and control
# groups within each major cell type from the integrated Imaging Mass
# Cytometry (IMC) Seurat object and visualizes the results as a heatmap.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Heatmap showing differential marker expression across cell types
#
# Notes:
# Analysis parameters are currently configured for the Precentral gyrus dataset.
# For analyses of other brain regions, please refer to "IMC_analysis_paramaters.txt"
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================


## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load previously saved object
```{r}
combined_seurat <- readRDS(
   file = file.path("rds/combined_seurat_PrecentralGyrus.rds")
)
```

## Markers info
```{r}
Markers_astro <- c("Cell-141Pr-AQP4", "Cell-158Gd-GFAP")
Markers_immune <- c("Cell-161Dy-CX3CR1", "Cell-164Dy-TMEM119","Cell-160Gd-IBA1","Cell-163Dy-CD4", "Cell-155Gd-FoxP3", "Cell-167Er-CD8")
Markers_neuron <- c("Cell-151Eu-MAP2","Cell-149Sm-C9orf72", "Cell-169Tm-ChAT","Cell-166Er-GAD67")
Markers_others <- c("Cell-175Lu-TDP43",
"Cell-154Sm-pTDP43",
"Cell-168Er-FUS",
"Cell-152Sm-ATXN2",
"Cell-145Nd-UBQLN2",
"Cell-170Er-APOE",
"Cell-176Yb-TBK1",
"Cell-162Dy-pTBK1",
"Cell-147Sm-Glutaminase",
"Cell-143Nd-SQSTM1",
"Cell-156Gd-CH24H",
"Cell-153Eu-ORF1",
"Cell-159Tb-IL6ST",
"Cell-165Ho-PD1",
"Cell-150Nd-PDL1",
"Cell-172Yb-PDL2",
"Cell-171Yb-TREM2",
"Cell-174Yb-MHCclassII"
)

Markers <- c(Markers_astro, Markers_immune, Markers_neuron, Markers_others)

Marker_label <- c(
  "Cell-141Pr-AQP4" = "AQP4",
  "Cell-143Nd-SQSTM1" = "SQSTM1",
  "Cell-145Nd-UBQLN2" = "UBQLN2",
  "Cell-147Sm-Glutaminase" = "GLS",
  "Cell-149Sm-C9orf72" = "NeuN",
  "Cell-150Nd-PDL1" = "PDL1",
  "Cell-151Eu-MAP2" = "MAP2",
  "Cell-152Sm-ATXN2" = "ATXN2",
  "Cell-153Eu-ORF1" = "ORF1",
  "Cell-154Sm-pTDP43" = "pTDP43",
  "Cell-155Gd-FoxP3" = "FOXP3",
  "Cell-156Gd-CH24H" = "CH24H",
  "Cell-158Gd-GFAP" = "GFAP",
  "Cell-159Tb-IL6ST" = "IL6ST",
  "Cell-160Gd-IBA1" = "IBA1",
  "Cell-161Dy-CX3CR1" = "CX3CR1",
  "Cell-162Dy-pTBK1" = "pTBK1",
  "Cell-163Dy-CD4" = "CD4",
  "Cell-164Dy-TMEM119" = "TMEM119",
  "Cell-165Ho-PD1" = "PD1",
  "Cell-166Er-GAD67" = "GAD67",
  "Cell-167Er-CD8" = "CD8",
  "Cell-168Er-FUS" = "FUS",
  "Cell-169Tm-ChAT" = "ChAT",
  "Cell-170Er-APOE" = "APOE",
  "Cell-171Yb-TREM2" = "TREM2",
  "Cell-172Yb-PDL2" = "PDL2",
  "Cell-174Yb-MHCclassII" = "MHCII",
  "Cell-175Lu-TDP43" = "TDP43",
  "Cell-176Yb-TBK1" = "TBK1"
)

```

# Heatmap drawing
```{r}

celltypes_use <- c("Astrocyte", "Immune", "Excitatory", "Inhibitory", "Others")
celltypes_use <- intersect(celltypes_use, unique(combined_seurat$Celltype))

selected_markers <- c(Markers_others)

compute_stats_one <- function(seu, celltype, markers,
                              assay = "Protein", slot = "counts") {
  obj <- subset(
    seu,
    subset = Celltype == celltype & major_group %in% c("ALS","Control")
  )
  
   mat <- GetAssayData(obj, assay = assay, slot = slot)
  
   als_cells <- WhichCells(obj, expression = major_group == "ALS")
   ctrl_cells <- WhichCells(obj, expression = major_group == "Control")
   
   mean_als    <- rowMeans(mat[, als_cells, drop = FALSE])
   mean_ctrl   <- rowMeans(mat[, ctrl_cells, drop = FALSE])
   mean_all    <- rowMeans(mat)

   sd_als  <- apply(mat[, als_cells, drop = FALSE], 1, sd)
   sd_ctrl <- apply(mat[, ctrl_cells, drop = FALSE], 1, sd)

   n_als  <- length(als_cells)
   n_ctrl <- length(ctrl_cells)
   
   s_pooled <- sqrt(
    ((n_als - 1)  * sd_als^2 +
       (n_ctrl - 1) * sd_ctrl^2) /
      (n_als + n_ctrl - 2)
   )
 
   cohens_d <- (mean_als - mean_ctrl) / s_pooled
   log2fc   <- log2((mean_als + 1) / (mean_ctrl + 1)) 
  
  ## Wilcoxon 
  p_values <- apply(mat, 1, function(x) {
    wilcox.test(x[als_cells], x[ctrl_cells])$p.value
  })
  p_adj   <- p.adjust(p_values, method = "BH")
  log10_p <- -log10(p_adj)

  ## q = |d| * -log10(FDR)
  q_values <- abs(cohens_d) * log10_p
  signif_label <- ifelse(q_values > 10, "*", "")
  
   df <- data.frame(
    Gene        = rownames(mat),
    Celltype    = celltype,
    log2FC      = log2fc,
    Ave_ALS     = mean_als,
    Ave_Control = mean_ctrl,
    Ave_all     = mean_all,
    CohensD     = cohens_d,
    p_val       = p_values,
    p_val_adj   = p_adj,
    log10_p     = log10_p,
    q_value     = q_values,
    Significance = signif_label,
    stringsAsFactors = FALSE
  ) 
  
   df <- df[df$Gene %in% markers, , drop = FALSE]
  } 
  
  stats_list <- lapply(
  celltypes_use,
  compute_stats_one,
  seu     = combined_seurat,
  markers = selected_markers
)
  stats_all <- bind_rows(stats_list)
  
  
  
  marker_order <- selected_markers[selected_markers %in% stats_all$Gene]

stats_all$Gene <- factor(
  stats_all$Gene,
  levels = rev(marker_order)
)

stats_all$Celltype <- factor(
  stats_all$Celltype,
  levels = celltypes_use
)

max_d <- 1.5  

p <- ggplot(stats_all,
            aes(x = Celltype, y = Gene, fill = CohensD)) +
  geom_tile(color = "white", size = 1) +
  geom_text(aes(label = Significance),
            color = "black", size = 10, vjust = 0.8) +
  scale_fill_gradientn(
   colours = c("blue", "#F2F2F2", "white", "#F2F2F2","red", "red"), 
      values = scales::rescale(c(-2, -0.2, 0, 0.2, 1.5,2)),
   
    limits  = c(-max_d, max_d),
    oob     = scales::squish,
    name    = "Cohen's d\n(ALS − Control)"
  ) +
  scale_y_discrete(labels = Marker_label[rev(marker_order)]) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1,
                               size = 14, color = "black"),
    axis.text.y = element_text(size = 14, color = "black"),
    legend.position = "right"
  )

print(p)
 
```

