# =============================================================================
# IMC_subclustering.R
#
# Description:
# This script performs subclustering analysis of major cell populations.
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Subcluster-annotated Seurat objects for:
#   - Astrocytes
#   - Immune cells
#   - Excitatory neurons
#   - Inhibitory neurons
# - UMAP plots for each subclustered population
#
# Notes:
# Analysis parameters are currently configured for the Precentral gyrus dataset.
# For analyses of other brain regions, please refer to "IMC_analysis_paramaters.txt"
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================


## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load previously saved object
```{r}
combined_seurat <- readRDS(
   file = file.path("rds/combined_seurat_PrecentralGyrus.rds")
)
combined_seurat_Astro <- subset(combined_seurat, subset = Celltype == "Astrocyte")
combined_seurat_Immune <- subset(combined_seurat, subset = Celltype == "Immune")
combined_seurat_Exc <- subset(combined_seurat, subset = Celltype == "Excitatory")
combined_seurat_Inh <- subset(combined_seurat, subset = Celltype == "Inhibitory")
```


## Subclustering 
# Astro
```{r}
combined_seurat_Astro <- FindNeighbors(combined_seurat_Astro, reduction = "harmony", dims = 1:20) %>%
        FindClusters(resolution = 0.5) %>%
        RunUMAP(reduction = "harmony", dims = 1:20, min.dist = 0.1, spread = 1.2)
DimPlot(combined_seurat_Astro, label = T)
DimPlot(combined_seurat_Astro, label = T, group.by = "major_group")

cluster_name <- RenameIdents(combined_seurat_Astro, 
                   '0' = "Astro_1",
                    '1' = "Astro_2",
                    '2' = "Astro_3",
                    '3' = "Astro_4",
                    '4' = "Astro_5"
                    )
combined_seurat_Astro$Subcelltype = Idents(cluster_name)

p <- DimPlot(combined_seurat_Astro, 
           group.by= "Subcelltype",
           label=T, 
           label.size =6,
           pt.size = 0.3) +
  theme_classic() +
  xlim(-7,7) +
  ylim(-7,7) +
   scale_color_manual(values = c(
"Astro_1" = "#FF5A5A",   
"Astro_2" = "#FF7794",  
"Astro_3" = "#FF8FC2", 
"Astro_4" = "#FFB0D8", 
"Astro_5" = "#FFD4C2"
  )) +
 
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text = element_text(size = 20, color = "black")
  )

print(p)


```

# Immune
```{r}
combined_seurat_Immune <- FindNeighbors(combined_seurat_Immune, reduction = "harmony", dims = 1:20) %>%
        FindClusters(resolution = 0.3) %>%
        RunUMAP(reduction = "harmony", dims = 1:20, min.dist = 0.1, spread = 0.8)
DimPlot(combined_seurat_Immune, label = T)
DimPlot(combined_seurat_Immune, label = T, group.by = "major_group")

cluster_name <- RenameIdents(combined_seurat_Immune, 
                   '0' = "Immune_1",
                   '1' = "Immune_2",
                   '2' = "Immune_3"
                    )
combined_seurat_Immune$Subcelltype = Idents(cluster_name)

p <- DimPlot(combined_seurat_Immune, 
           group.by= "Subcelltype",
           label=T, 
           label.size =6,
           pt.size = 0.3) +
  theme_classic() +
  xlim(-9,9) +
  ylim(-9,9) +
   scale_color_manual(values = c(
"Immune_1" = "#FF7A3D",  
"Immune_2" = "#FFAA33",  
"Immune_3" = "#FFE066"   
  )) +
  
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text = element_text(size = 20, color = "black")
  )

print(p)

```

# Excitatory
```{r}
combined_seurat_Exc <- FindNeighbors(combined_seurat_Exc, reduction = "harmony", dims = 1:20) %>%
        FindClusters(resolution = 0.6) %>%
        RunUMAP(reduction = "harmony", dims = 1:20, min.dist = 0.05, spread = 1.5)
DimPlot(combined_seurat_Exc, label = T)
DimPlot(combined_seurat_Exc, label = T, group.by = "major_group")
DimPlot(combined_seurat_Exc, label = T, group.by = "group")

cluster_name <- RenameIdents(combined_seurat_Exc, 
                  '0' = "Exc_1",
                  '1' = "Exc_2",
                  '2' = "Exc_3",
                  '3' = "Exc_4",
                   '4' = "Exc_5"
                                )
combined_seurat_Exc$Subcelltype = Idents(cluster_name)

p <- DimPlot(combined_seurat_Exc, 
           group.by= "Subcelltype",
           label=T, 
           label.size =6,
           pt.size = 0.3) +
  theme_classic() +
  xlim(-9,9) +
  ylim(-9,9) +
   scale_color_manual(values = c(
Exc_1 = "#CFEAFF",  
Exc_2 = "#9BD3FF", 
Exc_3 = "#5CB6FF",   
Exc_4 = "#1F78FF",   
Exc_5 = "#004A99"
  )) +
  
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text = element_text(size = 20, color = "black")
  )

print(p)
```


# Inhibitory
```{r}
combined_seurat_Inh <- FindNeighbors(combined_seurat_Inh, reduction = "harmony", dims = 1:18) %>%
        FindClusters(resolution = 0.5) %>%
        RunUMAP(reduction = "harmony", dims = 1:18, min.dist = 0.1, spread = 1)
DimPlot(combined_seurat_Inh, label = T)
DimPlot(combined_seurat_Inh, label = T, group.by = "major_group")
DimPlot(combined_seurat_Inh, label = T, group.by = "group")

cluster_name <- RenameIdents(combined_seurat_Inh, 
                  '0' = "Inh_1",
                  '1' = "Inh_2"
               
                    )
combined_seurat_Inh$Subcelltype = Idents(cluster_name)

p <- DimPlot(combined_seurat_Inh, 
           group.by= "Subcelltype",
           label=T, 
           label.size =6,
           pt.size = 0.3) +
  theme_classic() +
  xlim(-7,7) +
  ylim(-7,7) +
   scale_color_manual(values = c(
"Inh_1" = "#A6E65A",
"Inh_2" = "#5CBF3A"
      )) +
  
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(5, "pt"),
    strip.text = element_text(size = 20, color = "black")
  )

print(p)

```

#Combine
```{r}

combined_seurat@meta.data[Cells(combined_seurat_Astro), "Subcelltype"] <-   as.character(combined_seurat_Astro$Subcelltype)

combined_seurat@meta.data[Cells(combined_seurat_Immune), "Subcelltype"] <-   as.character(combined_seurat_Immune$Subcelltype)

combined_seurat@meta.data[Cells(combined_seurat_Exc), "Subcelltype"] <-   as.character(combined_seurat_Exc$Subcelltype)

combined_seurat@meta.data[Cells(combined_seurat_Inh), "Subcelltype"] <-   as.character(combined_seurat_Inh$Subcelltype)

combined_seurat$Subcelltype <- as.character(combined_seurat$Subcelltype)

cells_to_update <- Cells(subset(combined_seurat, Celltype == "Others"))
combined_seurat@meta.data[cells_to_update, "Subcelltype"] <- "Others"

combined_seurat$Subcelltype <- factor(combined_seurat$Subcelltype)

combined_seurat$Subcelltype <- factor(
  combined_seurat$Subcelltype,
  levels = c(
    paste0("Astro_", 1:5),
    paste0("Immune_", 1:3),
    paste0("Exc_", 1:5),
    paste0("Inh_", 1:2),
    "Others"
  )
)

p <- DimPlot(
  combined_seurat,
  group.by = "Subcelltype",
  label = TRUE,
  label.size = 4,
  pt.size = 0.2
) +
  scale_color_manual(values = c(
    
  # ---- Astro ----
"Astro_1" = "#FF5A5A",   
"Astro_2" = "#FF7794",  
"Astro_3" = "#FF8FC2", 
"Astro_4" = "#FFB0D8", 
"Astro_5" = "#FFD4C2",  

  # ---- Immune ----
"Immune_1" = "#FF7A3D",  
"Immune_2" = "#FFAA33",  
"Immune_3" = "#FFE066",

  # ---- Exc ----
"Exc_1" = "#CFEAFF",  
"Exc_2" = "#9BD3FF", 
"Exc_3" = "#5CB6FF",   
"Exc_4" = "#1F78FF",   
"Exc_5" = "#004A99",  

  # ---- Inh ----
"Inh_1" = "#A6E65A",
"Inh_2" = "#5CBF3A",

  # ---- Others ----
"Others" = "#A3A500"
  )) +
  
  ggtitle("Subcelltype")

print(p)


```

## Save integrated Seurat object
```{r}
#saveRDS(
#  combined_seurat,
#  file = file.path(output_folder, "combined_seurat_PrecentralGyrus.rds")
#)

#saveRDS(
#  combined_seurat_Astro,
#  file = file.path(output_folder, "combined_seurat_Astro_PrecentralGyrus.rds")
#)

#saveRDS(
#  combined_seurat_Immune,
#  file = file.path(output_folder, "combined_seurat_Immune_PrecentralGyrus.rds")
#)

#saveRDS(
#  combined_seurat_Exc,
#  file = file.path(output_folder, "combined_seurat_Exc_PrecentralGyrus.rds")
#)

#saveRDS(
#  combined_seurat_Inh,
#  file = file.path(output_folder, "combined_seurat_Inh_PrecentralGyrus.rds")
#)

```


