# =============================================================================
# IMC_Seurat_clustering.R
#
# Description:
# This script performs single-cell protein clustering of Imaging Mass Cytometry
# (IMC) data using the Seurat framework. 
#
# Author:
# Momoka Hikosaka
# Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Preprocessed IMC single-cell expression matrices (*.csv)
#
# Output:
# - Integrated Seurat object
# - PCA, Harmony, and UMAP embeddings
# - Cluster annotations
# - Visualization figures
#
# Notes:
# Analysis parameters are currently configured for the Precentral gyrus dataset.
# For analyses of other brain regions, please refer to "IMC_analysis_paramaters.txt"
#
# License:
# Copyright (c) 2025 Momoka Hikosaka
# This code is provided for academic research purposes.
# =============================================================================


## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(harmony)
  library(tidyverse)
  library(ggplot2)
  library(viridis)
  library(patchwork)
  library(harmony)
  library(cowplot)
  library(tools)
  library(forcats) 
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Create Seurat objects from selected CSV files
```{r}
Data_path = "../preprocessed_csv"

Data_name = c("PrecentralGyrus_01_ALS",
              "PrecentralGyrus_02_ALS",
              "PrecentralGyrus_04_ALS_1",
              "PrecentralGyrus_04_ALS_2",
              "PrecentralGyrus_06_ALS",
              "PrecentralGyrus_08_AC",
              "PrecentralGyrus_09_AC",
              "PrecentralGyrus_11_AC",
              "PrecentralGyrus_12_AC"
              )

Name = c("ALS_1",
              "ALS_2",
              "ALS_4_1",
              "ALS_4_2",
              "ALS_6",
              "Control_8",
              "Control_9",
              "Control_11",
              "Control_12"
              )

Sample_name = c("ALS_1",
              "ALS_2",
              "ALS_4",
              "ALS_4",
              "ALS_6",
              "Control_8",
              "Control_9",
              "Control_11",
              "Control_12"
              )

Data_table = data.frame(Data_name, Name, Sample_name, stringsAsFactors = FALSE)

files <- lapply(1:nrow(Data_table), function(i) {
  list(
    path = paste0(Data_path,  "/", Data_table$Data_name[i], "_preprocessed.csv"),
    group = Data_table$Name[i],
    sample_name  = Data_table$Sample_name[i]  
  )
})

seurat_list <- list()

for (i in seq_along(files)) {
  
  temp <- read.csv(files[[i]]$path, header = TRUE)
   if ("Sel01" %in% colnames(temp)) {
    seurat_data <- temp[temp[["Sel01"]] %in% c(1, "1"), , drop = FALSE]
  } else {
    seurat_data <- temp   
  }

  seurat_data <- seurat_data[, 2:32]
  rownames(seurat_data) <- seurat_data$CellId
  seurat_data <- seurat_data[, 2:31]
 
  expression_matrix <- t(as.matrix(seurat_data))
  
  seurat_object <- CreateSeuratObject(counts = expression_matrix, assay = "Protein")
  DefaultAssay(seurat_object) <- "Protein"
  
  seurat_object$group <- files[[i]]$group
  seurat_object$Sample_name <- files[[i]]$sample_name
  
  seurat_object@assays$Protein$data <- as.matrix(seurat_object@assays$Protein$counts)
seurat_object <- ScaleData(seurat_object, assay = "Protein")

   seurat_list[[i]] <- seurat_object
 }



by_sample <- split(seurat_list, sapply(seurat_list, function(x) unique(x$Sample_name)))

for (sample in names(by_sample)) {
  lst <- by_sample[[sample]]
  for (j in seq_along(lst)) {
    lst[[j]]$run_id <- paste0(sample, "_run", j)
  }
  by_sample[[sample]] <- lst
}


seurat_per_sample <- lapply(by_sample, function(lst) {
  merged <- Reduce(function(x, y) merge(x, y), lst)
  if ("layers" %in% slotNames(merged[["Protein"]])) {
    merged[["Protein"]] <- JoinLayers(merged[["Protein"]])
  }
  merged
})

combined_seurat <- Reduce(function(x, y) merge(x, y), seurat_per_sample)

combined_seurat[["Protein"]] <- JoinLayers(combined_seurat[["Protein"]])


  
combined_seurat <- FindVariableFeatures(combined_seurat, assay = "Protein")
features <-VariableFeatures(combined_seurat)

Ex_marker <- c("Cell-165Ho-PD1",
"Cell-150Nd-PDL1",
"Cell-172Yb-PDL2",
"Cell-171Yb-TREM2",
"Cell-174Yb-MHCclassII") 
features<- setdiff(features, Ex_marker)
VariableFeatures(combined_seurat) <- features


#PCA
combined_seurat <- RunPCA(combined_seurat, assay = "Protein", features = VariableFeatures(combined_seurat))

## Harmony integration
combined_seurat <- RunHarmony(
  object = combined_seurat,
   group.by.vars = c("Sample_name", "run_id")
)

```

## UMAP clustring
Note: UMAP embeddings may show slight differences in layout across runs owing to stochastic optimization. For exact reproduction of previously generated embeddings, the saved RDS object should be loaded instead of rerunning the workflow.
```{r}
combined_seurat <- FindNeighbors(combined_seurat,
                                 reduction = "harmony",
                                 dims      = 1:20,
                                 k.param   = 70)  

combined_seurat <- FindClusters(combined_seurat, resolution = 2)

combined_seurat <- RunUMAP(combined_seurat,
                           reduction = "harmony",
                           dims      = 1:20,
                           min.dist  = 0.1,
                           spread    = 2,
                           n.neighbors = 70) 

DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, group.by = "run_id")
DimPlot(combined_seurat, label = T, group.by = "group")
DimPlot(combined_seurat, label = T, group.by = "Sample_name")
DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, split.by = "group")
DimPlot(combined_seurat, label = T, split.by = "Sample_name")
```


## Heatmap to check marker expression patterns
```{r}
DefaultLayer(combined_seurat[["Protein"]]) <- "data"

markers_astro <- c(
  "Cell-141Pr-AQP4",
  "Cell-158Gd-GFAP"
)

markers_immune <- c(
  "Cell-161Dy-CX3CR1",
  "Cell-164Dy-TMEM119",
  "Cell-160Gd-IBA1",
  "Cell-163Dy-CD4",
  "Cell-155Gd-FoxP3",
  "Cell-167Er-CD8"
)

markers_neuron <- c(
  "Cell-151Eu-MAP2",
  "Cell-149Sm-C9orf72",
  "Cell-169Tm-ChAT",
  "Cell-166Er-GAD67"
)

markers_other <- c(
  "Cell-175Lu-TDP43",
  "Cell-154Sm-pTDP43",
  "Cell-168Er-FUS",
  "Cell-152Sm-ATXN2",
  "Cell-145Nd-UBQLN2",
  "Cell-170Er-APOE",
  "Cell-176Yb-TBK1",
  "Cell-162Dy-pTBK1",
  "Cell-147Sm-Glutaminase",
  "Cell-143Nd-SQSTM1",
  "Cell-156Gd-CH24H",
  "Cell-153Eu-ORF1",
  "Cell-159Tb-IL6ST",
  "Cell-165Ho-PD1",
  "Cell-150Nd-PDL1",
  "Cell-172Yb-PDL2",
  "Cell-171Yb-TREM2",
  "Cell-174Yb-MHCclassII"
)

markers <- c(markers_astro, markers_immune, markers_neuron, markers_other)

marker_label <- c(
  "Cell-141Pr-AQP4" = "AQP4",
  "Cell-143Nd-SQSTM1" = "SQSTM1",
  "Cell-145Nd-UBQLN2" = "UBQLN2",
  "Cell-147Sm-Glutaminase" = "GLS",
  "Cell-149Sm-C9orf72" = "NeuN",
  "Cell-150Nd-PDL1" = "PDL1",
  "Cell-151Eu-MAP2" = "MAP2",
  "Cell-152Sm-ATXN2" = "ATXN2",
  "Cell-153Eu-ORF1" = "ORF1",
  "Cell-154Sm-pTDP43" = "pTDP43",
  "Cell-155Gd-FoxP3" = "FOXP3",
  "Cell-156Gd-CH24H" = "CH24H",
  "Cell-158Gd-GFAP" = "GFAP",
  "Cell-159Tb-IL6ST" = "IL6ST",
  "Cell-160Gd-IBA1" = "IBA1",
  "Cell-161Dy-CX3CR1" = "CX3CR1",
  "Cell-162Dy-pTBK1" = "pTBK1",
  "Cell-163Dy-CD4" = "CD4",
  "Cell-164Dy-TMEM119" = "TMEM119",
  "Cell-165Ho-PD1" = "PD1",
  "Cell-166Er-GAD67" = "GAD67",
  "Cell-167Er-CD8" = "CD8",
  "Cell-168Er-FUS" = "FUS",
  "Cell-169Tm-ChAT" = "ChAT",
  "Cell-170Er-APOE" = "APOE",
  "Cell-171Yb-TREM2" = "TREM2",
  "Cell-172Yb-PDL2" = "PDL2",
  "Cell-174Yb-MHCclassII" = "MHCII",
  "Cell-175Lu-TDP43" = "TDP43",
  "Cell-176Yb-TBK1" = "TBK1"
)

heatmap_data <- DotPlot(combined_seurat, features = markers)$data
heatmap_data$id <- factor(
  heatmap_data$id,
  levels = sort(as.numeric(unique(heatmap_data$id)))
)

p <- ggplot(heatmap_data, aes(x = features.plot, y = fct_rev(id))) +
  geom_tile(colour = "white", aes(fill = avg.exp.scaled)) +
  scale_fill_gradientn(
    colours = c("lightyellow3", "lightyellow3", "lightyellow", "orange", "red"),
    values = c(0, 0.2, 0.5, 0.75, 1),
    name = "Normalized Expression"
  ) +
  scale_x_discrete(labels = marker_label) +
  ggtitle("Heatmap of marker expression") +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, face = "italic"),
    legend.position = "bottom"
  )

print(p)
```

## UMAP annotation
```{r}
cell_type <- RenameIdents(
  combined_seurat,
  "0" = "Excitatory",
  "1" = "Others",
  "2" = "Others",
  "3" = "Astrocyte",
  "4" = "Excitatory",
  "5" = "Excitatory",
  "6" = "Others",
  "7" = "Excitatory",
  "8" = "Astrocyte",
  "9" = "Inhibitory",
  "10" = "Astrocyte",
  "11" = "Others",
  "12" = "Immune",
  "13" = "Immune",
  "14" = "Astrocyte",
  "15" = "Immune"
)

combined_seurat$Celltype <- Idents(cell_type)

p <- DimPlot(
  combined_seurat,
  group.by = "Celltype",
  pt.size = 0.3,
  label = TRUE,
  label.size = 6
) +
  ggtitle("Major cell type") +
  scale_color_manual(
    values = c(
      "Astrocyte" = "#F8766D",
      "Immune" = "#FFD700",
      "Excitatory" = "#00B0F6",
      "Inhibitory" = "#00BF7D",
      "Others" = "#A3A500"
    )
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", linewidth = 0.5),
    axis.ticks.length = unit(5, "pt")
  )

print(p)
 
```

## Group metadata
```{r}
combined_seurat$major_group <- ifelse(
  grepl("^ALS", combined_seurat$group),
  "ALS",
  "Control"
)

combined_seurat_ALS <- subset(combined_seurat, subset = major_group == "ALS")
combined_seurat_control <- subset(combined_seurat, subset = major_group == "Control")
```

## Save integrated Seurat object
```{r}
#saveRDS(
#  combined_seurat,
#  file = file.path(output_folder, "combined_seurat_PrecentralGyrus.rds")
#)
```

## Load previously saved object
Note: Use this section when you want to continue the analysis from a previously saved Seurat object instead of rerunning the full preprocessing and integration workflow.
```{r}
combined_seurat <- readRDS(
 file = file.path("rds/combined_seurat_PrecentralGyrus.rds")
)
```

## UMAP clustring
```{r}

#set.seed(1234)

#combined_seurat <- FindNeighbors(combined_seurat,
#                                 reduction = "harmony",
#                                 dims      = 1:20,
#                                 k.param   = 70)  

#combined_seurat <- FindClusters(combined_seurat, resolution = 2)

#combined_seurat <- RunUMAP(combined_seurat,
#                           reduction = "harmony",
#                           dims      = 1:20,
#                           min.dist  = 0.1,
#                           spread    = 2,
#                           n.neighbors = 70) 

Idents(combined_seurat) =  "seurat_clusters"
DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, group.by = "run_id")
DimPlot(combined_seurat, label = T, group.by = "group")
DimPlot(combined_seurat, label = T, group.by = "Sample_name")
DimPlot(combined_seurat, label = T)
DimPlot(combined_seurat, label = T, split.by = "group")
DimPlot(combined_seurat, label = T, split.by = "Sample_name")


p <- DimPlot(
  combined_seurat,
  group.by = "Celltype",
  pt.size = 0.3,
  label = TRUE,
  label.size = 6
) +
  ggtitle("Major cell type") +
  scale_color_manual(
    values = c(
      "Astrocyte" = "#F8766D",
      "Immune" = "#FFD700",
      "Excitatory" = "#00B0F6",
      "Inhibitory" = "#00BF7D",
      "Others" = "#A3A500"
    )
  ) +
  theme_classic() +
  theme(
    axis.text = element_text(size = 20, color = "black"),
    axis.title = element_text(size = 20, color = "black"),
    legend.position = "right",
    axis.line = element_line(color = "black", linewidth = 0.5),
    axis.ticks.length = unit(5, "pt")
  )

print(p)
 
```

