# =============================================================================
# scRNAseq_expression.R
#
# Description:
# This script loads a integrated Seurat object and
# performs gene expression visualization and statistical analysis.
#
# Author:
# Md. Shamim Hossain, Momoka Hikosaka, Suthinee Attachaipanich
# Institute of Rheology Functions of Food, Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Input:
# - Integrated Seurat object
#
# Output:
# - Violin plots of gene expression per cell type
# - Dot plots summarizing differential expression
# - Statistical test results (Wilcoxon test)
#
# License:
# Copyright (c) 2025  Md. Shamim Hossain, Momoka Hikosaka, Suthinee Attachaipanich
# This code is provided for academic research purposes.
# =============================================================================

## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(Seurat)
  library(patchwork)
  library(tidyverse)
  library(gtools)
  library(harmony)
  library(ggrepel)
  library(ggplot2)
})
```

## Output directory
```{r}
output_folder <- "output"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}
```

## Load previously saved object
```{r}
Int <- readRDS(
  file = file.path("rds/Pineda.Int.rds")
)

table(Int$CellClass)
table(Int$CellType)

Idents(Int) = Int$CellType
DimPlot(Int, label =T)
```

# Violin plot
```{r}
celltypes <- c("Astro","Micro")
#"Endo","Fibro","Micro","Mural","Oligo","OPC","T_Cell","Oligo","L5_L6"

genes_of_interest <- c("GLS","GLS2","TP53","MYC")
#"ATM","CHEK2","BBC3","NFE2L2","BAX","CDKN1A","MAPK14","MAPK11","MAP3K20","MAPK8","MAPK9","MAPK10","EIF2AK4","EIF2A"
#"MAPK14","MAPK11"P38
#"MAP3K20"ZAK
#"MAPK8"JNK1
#"MAPK9"JNK2
#"MAPK10" JNK3
#"EIF2AK4","EIF2A"

results_list <- list()

for (ct in celltypes) {

  ct_subset <- subset(Int, subset = CellType == ct)

  for (gene in genes_of_interest) {

    if (!(gene %in% rownames(ct_subset))) {
      message(gene, " not found in dataset (", ct, ")")
      next
    }

    expr_data <- GetAssayData(ct_subset, slot = "counts")[gene, ]
    cells_to_keep <- colnames(ct_subset)[expr_data > 0]

    if (length(cells_to_keep) == 0) {
      message("No expressing cells for ", gene, " in ", ct)
      next
    }

    filtered_subset <- subset(ct_subset, cells = cells_to_keep)

    p <- VlnPlot(filtered_subset,
                 features = gene,
                 group.by = "Condition",
                 pt.size = 0.05,
                 cols = c("red", "blue"),
                 adjust = 2) +
      coord_cartesian(ylim = c(0, 3.5)) +
      ggtitle(paste(ct, gene, sep=" : ")) +
      NoLegend() +
      geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")

    meta <- filtered_subset@meta.data
    expr_vals <- GetAssayData(filtered_subset, slot = "data")[gene, ]  

  
    als_cells <- rownames(meta)[meta$Condition == "ALS"]
    pn_cells  <- rownames(meta)[meta$Condition == "PN"]

    x <- expr_vals[als_cells]
    y <- expr_vals[pn_cells]

   
    min_cells_per_group <- 3

    if (length(x) >= min_cells_per_group && length(y) >= min_cells_per_group) {

      wt <- tryCatch(
        wilcox.test(x, y, exact = FALSE),
        error = function(e) NULL
      )

      if (!is.null(wt) && !is.na(wt$p.value)) {

      mean_ALS <- mean(x, na.rm = TRUE)
mean_PN  <- mean(y, na.rm = TRUE)

results_list[[length(results_list) + 1]] <- data.frame(
  CellType = ct,
  Gene     = gene,
  n_ALS    = length(x),
  n_PN     = length(y),
  Mean_ALS = mean_ALS,
  Mean_PN  = mean_PN,
  W_stat   = unname(wt$statistic),
  p_val    = wt$p.value,
  stringsAsFactors = FALSE
)
       

      } else {
        results_list[[length(results_list) + 1]] <- data.frame(
          CellType = ct, Gene = gene,
          n_ALS = length(x), n_PN = length(y),
          W_stat = NA_real_, p_val = NA_real_,
          stringsAsFactors = FALSE
        )
      }

    } else {
      results_list[[length(results_list) + 1]] <- data.frame(
        CellType = ct, Gene = gene,
        n_ALS = length(x), n_PN = length(y),
        W_stat = NA_real_, p_val = NA_real_,
        stringsAsFactors = FALSE
      )
    }

 
    print(p)

    file_name <- paste0("Vlnplot_filter_", ct, "_", gene, ".pdf")
    ggsave(filename = file.path(output_folder, file_name),
           plot = p, width = 3.5, height = 6, dpi = 300)
  }
}


results_table <- bind_rows(results_list)

write.csv(results_table,
          file = file.path(output_folder, "Wilcoxon_results.csv"),
          row.names = FALSE)


```

# Dotplot
```{r}
library(dplyr)
library(ggplot2)
library(Seurat)


celltypes <- c("Astro","Micro","Oligo","L2_L3","L5_L6","PV","SOM")
genes_of_interest <- c("GRN","LAMP1","LAMP2","CTSD","SQSTM1","TFEB",paste0("blank", 1:10) )
#"GRN","LAMP1","LAMP2","CTSD","SQSTM1","TFEB"
#"VIM","NES","CSPG4","VCAN","TNC","COL1A1","COL4A1","FN1","C3","SERPING1","MMP2","MMP9","CD44","S100B", paste0("blank", 1:2) 
#"TGFB1","TGFB2","TGFB3","TGFBR1","TGFBR2","TGFBR3","SMAD2","SMAD3","SMAD4","SMAD7", paste0("blank", 1:6) 
#"IL6","IL6R","IL6ST","JAK1","JAK2","JAK3","TYK2","STAT1","STAT2","STAT3", paste0("blank", 1:6)

cond1 <- "ALS"
cond2 <- "PN"

plot_list <- list()

for (ct in celltypes) {

  obj_ct <- subset(Int, subset = CellType == ct)
  meta_all <- obj_ct@meta.data

   if (!all(c(cond1, cond2) %in% unique(meta_all$Condition))) next

 for (gene in genes_of_interest) {

    if (!(gene %in% rownames(obj_ct))) next

   
    expr_all <- GetAssayData(obj_ct, slot = "data")[gene, ]
    pct_expressed <- mean(expr_all > 0) * 100

   
    cells_to_keep <- colnames(obj_ct)[expr_all > 0]
    if (length(cells_to_keep) == 0) next

    filtered_subset <- subset(obj_ct, cells = cells_to_keep)
    meta_f <- filtered_subset@meta.data
    if (!all(c(cond1, cond2) %in% unique(meta_f$Condition))) next

    expr_f <- GetAssayData(filtered_subset, slot = "data")[gene, ]

    cond1_cells <- rownames(meta_f)[meta_f$Condition == cond1]
    cond2_cells <- rownames(meta_f)[meta_f$Condition == cond2]
    if (length(cond1_cells) < 2 || length(cond2_cells) < 2) next

    mean1 <- mean(expr_f[cond1_cells], na.rm = TRUE)
    mean2 <- mean(expr_f[cond2_cells], na.rm = TRUE)
    log2fc <- mean1 - mean2

    t_res <- tryCatch(
      t.test(expr_f[cond1_cells], expr_f[cond2_cells]),
      error = function(e) NULL
    )
    p_val <- if (!is.null(t_res)) t_res$p.value else NA_real_

    plot_list[[paste(ct, gene, sep="|")]] <- data.frame(
      CellType = ct,
      Gene = gene,
      pct_expressed = pct_expressed,
      log2fc = log2fc,
      p_val = p_val
    )
  }
}

plot_df <- bind_rows(plot_list)

blank_df <- expand.grid(
  CellType = celltypes,
  Gene = paste0("blank", 1:10)
)

blank_df$pct_expressed <- NA
blank_df$log2fc <- NA
blank_df$star <- ""

plot_df <- bind_rows(plot_df, blank_df)

p_to_star <- function(p) {
  if (is.na(p)) return("")
 
  if (p < 0.001)  return("*")
  return("")
}
plot_df$star <- vapply(plot_df$p_val, p_to_star, character(1))



plot_df$CellType <- factor(plot_df$CellType, levels = rev(celltypes))
plot_df$Gene <- factor(plot_df$Gene, levels = genes_of_interest)

plot_df <- plot_df %>%
  group_by(CellType) %>%
  mutate(star_y = as.numeric(CellType) -0.12 ) %>% #+ 0.18
  ungroup()

p <- ggplot(plot_df, aes(x = Gene, y = CellType)) +
  geom_point(aes(size = pct_expressed, color = log2fc)) +
   geom_text(aes(y = star_y, label = star), size = 6) +
  scale_size_area(
    max_size = 14,
    breaks = c(20, 40, 60),
    limits = c(0, 100),
    name = "% expressed"
  ) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red",   midpoint = 0,
  limits = c(-0.3, 0.3),
  oob = scales::squish) +
  theme_classic() +
  labs(
    title = "", #paste0(cond1, " - ", cond2, "  (color = log2FC from Violin, size = % expressed)"),
    size = "% expressed",
    color = "log2FC (ALS - PN)\n(nonzero cells)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    axis.title.y = element_text(size = 10),
    axis.title.x = element_text(size = 10),
    axis.text.x  = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
    axis.text.y  = element_text(size = 10),
    legend.title = element_text(size = 6),
    legend.text  = element_text(size = 6),
    plot.title   = element_text(size = 12),
    axis.line = element_line(color = "black", linewidth = 0.5),
  ) +
  
  guides(
      size = guide_legend(override.aes = list(
      shape = 21,        
      color = "black",   
      fill  = "white",   
      stroke = 0.5      
    ) ),
 color = guide_colorbar(barheight = 3, barwidth = 0.6)
)+
  scale_x_discrete(labels = function(x) ifelse(grepl("blank", x), "", x))

print(p)
```



