# =============================================================================
# scRNAseq_Seurat.R
#
# Description:
# This script performs Seurat-based integration and clustering analysis of
# publicly available single-cell RNA-seq data from ALS and control samples.
#
# Author:
# Md. Shamim Hossain, Momoka Hikosaka, Suthinee Attachaipanich
# Institute of Rheology Functions of Food, Kyoto University
#
# Associated publication:
# Hikosaka et al. in review
# "Glial metabolic states in human sporadic amyotrophic lateral sclerosis
# uncovered by spatial single-cell proteomics."
#
# Public dataset:
# Pineda et al., Cell, 2024
#
# Input:
# - count matrix file: counts_fil.mtx
# - cell metadata file: col_metadata.tsv
# - gene metadata file: row_metadata.tsv
#
# Output:
# - Integrated Seurat object
# - Harmony-corrected low-dimensional embedding
# - UMAP visualization
# - Cluster assignments
#
# License:
# Copyright (c) 2025  Md. Shamim Hossain, Momoka Hikosaka, Suthinee Attachaipanich
# This code is provided for academic research purposes.
# =============================================================================


## Environment cleanup
```{r}
rm(list = ls())  
```

## Load required packages
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(Seurat)
  library(patchwork)
  library(tidyverse)
  library(gtools)
  library(harmony)
  library(ggrepel)
  library(ggplot2)
})
```

```{r}
ProjectName <- c(
  "ALS101","ALS102","ALS103","ALS104","ALS106","ALS108","ALS110","ALS111",
  "ALS112","ALS116","ALS117","ALS118","ALS120","ALS122","ALS124","ALS126",
  "ALS127","Control301","Control302","Control303","Control304","Control306",
  "Control307","Control309","Control311","Control317","Control318","Control319",
  "Control322","Control323","Control324","Control325","Control328"
)

DirName <- c(
  "101_ALS","102_ALS","103_ALS","104_ALS","106_ALS","108_ALS","110_ALS","111_ALS",
  "112_ALS","116_ALS","117_ALS","118_ALS","120_ALS","122_ALS","124_ALS","126_ALS",
  "127_ALS","301_Control","302_Control","303_Control","304_Control","306_Control",
  "307_Control","309_Control","311_Control","317_Control","318_Control","319_Control",
  "322_Control","323_Control","324_Control","325_Control","328_Control"
)

condition <- c(rep("ALS", 17), rep("Control", 16))

sample_table <- data.frame(
  project_name = ProjectName,
  dir_name = DirName,
  condition = condition,
  stringsAsFactors = FALSE
)

base_dir <- "RNAseq_PublicData/Pineda_2024_Cell"

RNA <- vector("list", nrow(sample_table))

for (i in seq_len(nrow(sample_table))) {
  message("Creating Seurat object: ", i, " / ", nrow(sample_table),
          " (", sample_table$project_name[i], ")")

  sample_dir <- file.path(base_dir, sample_table$dir_name[i])

  metadata <- read.delim(file.path(sample_dir, "col_metadata.tsv"))
  features <- read.delim(file.path(sample_dir, "row_metadata.tsv"))
  Mtx1 <- Matrix::readMM(file.path(sample_dir, "counts_fil.mtx"))

  features$Gene[grepl("N/A", features$Gene)] <- NA
  features$NewGene <- NA

  rownames(features) <- features$ENSEMBL
  features[is.na(features$Gene), "NewGene"] <- features$ENSEMBL[is.na(features$Gene)]
  features[!is.na(features$Gene), "NewGene"] <- features$Gene[!is.na(features$Gene)]

  tmp1 <- features[!is.na(features$Gene), ]
  duplicated_genes <- tmp1$Gene[duplicated(tmp1$Gene)]
  tmp1 <- dplyr::filter(features, Gene %in% duplicated_genes)
  features[rownames(tmp1), "NewGene"] <- features[rownames(tmp1), "ENSEMBL"]

  colnames(Mtx1) <- metadata$Barcode
  rownames(Mtx1) <- features$NewGene
  rownames(metadata) <- metadata$Barcode

  RNA[[i]] <- CreateSeuratObject(
    counts = Mtx1,
    project = sample_table$project_name[i],
    min.cells = 3,
    min.features = 200,
    meta.data = metadata
  )

  RNA[[i]]$condition <- sample_table$condition[i]
  RNA[[i]][["percent.mt"]] <- PercentageFeatureSet(RNA[[i]], pattern = "^MT-")
  RNA[[i]] <- subset(RNA[[i]], subset = percent.mt <= 10)
}

Int <- merge(
  x = RNA[[1]],
  y = RNA[2:length(RNA)],
  merge.data = TRUE
)

Int[["RNA"]] <- JoinLayers(Int[["RNA"]])

Int <- Int %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA(verbose = FALSE)

Int <- RunHarmony(Int, group.by.vars = "orig.ident")
Int <- RunUMAP(Int, reduction = "harmony", dims = 1:30, min.dist = 0.3, spread = 1)
Int <- FindNeighbors(Int, reduction = "harmony", dims = 1:30) %>%
  FindClusters(resolution = 0.3)

# saveRDS(Int, file.path(base_dir, "Pineda.Int.rds"))

```
